---
title: "Razor Clam Co-management Staffing Model — Imputation Error for Long Beach & Twin Harbors (2026-27 planning)"
author: "WDFW Coastal Shellfish Unit"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    df_print: paged
---

<!--
  PURPOSE
  Model how the uncertainty (coefficient of variation, CV) of the SEASON effort and
  harvest estimates for Long Beach (LB) and Twin Harbors (TH) changes as a function of
  how many of their open beach-days are directly counted vs imputed from a single
  northern donor beach (Copalis OR Mocrocks — never both, because the 2-on/2-off
  co-management calendar means the two are never state-counted on the same day, so the
  joint donor of section 4.5 cannot be built).

  This is a decision tool for the 2026-27 field-staffing plan. It expands the v10
  methodology document's Figure 18 / Table 21 in four ways the document does not:
    (1) it resolves the answer by DONOR (Copalis vs Mocrocks vs the unavailable joint);
    (2) it splits the correction-factor error by light condition (daylight vs evening),
        which is the "model daylight vs night differently" question;
    (3) it stratifies by season part (fall / winter / spring); and
    (4) it translates the answer into crew-days for the actual 2026-27-shaped season.

  STRUCTURE follows 20260710-razor-clam-harvest-figures-pipeline.Rmd (load.lib loop,
  my_theme, dated 03_analyses/ output tree, save_plot with dpi=300). It reads ONLY the
  tidy CSVs in 02_data/consolidated/. set.seed and n_rep govern all Monte Carlo; keep
  them fixed for reproducible bands.

  NOTE ON EXECUTION: this document was authored without an R runtime available. Run it
  with rmarkdown::render() in RStudio. Columns are accessed by their exact raw CSV header
  (df[["header"]]) rather than janitor::clean_names(), so the parameter extraction does
  not depend on which name-cleaner is active.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r reset}
rm(list = ls())
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 0: TOGGLES
# ═══════════════════════════════════════════════════════════════════════════════

```{r toggles}
save_fig        <- TRUE        # master ggsave gate (set FALSE for fast iteration)
n_rep           <- 3000L       # Monte Carlo replicates per scenario
n_grid          <- 16L         # number of staffing levels swept (days counted)
set.seed(42)                   # reproducibility for ALL resampling / simulation
planning_season <- "2025-26"   # the most recent complete season, used as the 2026-27 skeleton

# --- error-model parameters that are NOT in a single consolidated CSV cell -------------
# Sourced from 2026-Review-...-v10.docx; expose them here so assumptions are visible.
SIG_CPUE_TRANSFER <- 0.448   # sd of log(donor CPUE / target CPUE) on paired days (v10 sec 4.5)
SIG_CPUE_SEASON   <- 0.137   # irreducible SEASON-level component of the CPUE-transfer error (v10 sec 11.1)
SIG_INT_DAY       <- 0.060   # per-counted-day interview (digging-fraction) log-sd on EFFORT (calibrated; least-constrained)
SIG_CPUE_MEAS_DAY <- 0.170   # per-counted-day measured-CPUE sampling log-sd on HARVEST (calibrated)
POOLED_IMP_1DAY   <- 0.541   # pooled single-day imputation log-sd (denominator for the season-part & aggregation shapes)
# SCOPE: this model uses a binary counted-vs-imputed state. Long Beach's intermediate
# "walk one half, scale the other section" option (section-ratio rule, log-sd ~0.281,
# v10 sec 5.4) is a cheaper tier that sits between a full count (0) and a full cross-beach
# borrow (~0.55-0.64). It is flagged as a future extension, not modelled here.

# Precision targets (external standards adopted in v10 sec 11.1)
CV_DESIGN <- 0.20   # creel-survey design target (Lester et al. 1991; McCormick & Meyer 2017)
CV_MRIP   <- 0.30   # NOAA MRIP publication threshold (PSE <= 30)
CV_TIGHT  <- 0.15   # a stricter internal target, shown for context
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 1: PACKAGES, THEME, HELPERS
# ═══════════════════════════════════════════════════════════════════════════════

```{r packages}
load.lib <- c("tidyverse", "readr", "janitor", "scales", "here", "patchwork")
install.lib <- load.lib[!load.lib %in% installed.packages()]
for (lib in install.lib) install.packages(lib, dependencies = TRUE)
sapply(load.lib, require, character.only = TRUE)

## ggplot theme (from the figure pipeline)
my_theme <- theme(text             = element_text(size = 14, color = "black"),
                  panel.background = element_blank(),
                  panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
                  panel.grid.minor = element_blank(),
                  axis.text        = element_text(size = 13, color = "black"),
                  axis.ticks       = element_line(color = "black"),
                  panel.border     = element_rect(color = "black", fill = NA, linewidth = 1.0),
                  strip.background = element_rect(fill = "grey95", color = "black"),
                  strip.text       = element_text(size = 13, face = "bold"),
                  legend.key       = element_blank(),
                  legend.position  = "bottom",
                  plot.title       = element_text(face = "bold"),
                  plot.margin      = margin(10, 20, 10, 10))

# small helper: numeric coercion that tolerates blank/character cells
num <- function(x) suppressWarnings(as.numeric(x))
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 2: OUTPUT FOLDER
# ═══════════════════════════════════════════════════════════════════════════════

```{r out_folder}
current_date <- format(Sys.Date(), "%Y%m%d")
# anchor at the repo root (here::here) so outputs land in 03_analyses/ regardless of the
# knit working directory (rmarkdown::render sets it to the .Rmd's folder, i.e. 01_code/).
folder_name  <- here::here("03_analyses", paste0(current_date, "-comanagement-staffing"))
subfolders   <- c("cv_curves", "cf_model", "season_part", "decision", "crew_days",
                  "variance", "validation", "tables")
for (sf in subfolders)
  dir.create(file.path(folder_name, sf), recursive = TRUE, showWarnings = FALSE)
cat("Output folder:", folder_name, "\n")

save_plot <- function(p, sub, name, w = 9, h = 6) {
  if (save_fig)
    ggsave(file.path(folder_name, sub, name), plot = p, width = w, height = h, dpi = 300)
}

# beach palette (matches beach_metadata.csv / the figure pipeline)
BEACH_COL <- c("Long Beach" = "#2166AC", "Twin Harbors" = "#66BD63",
               "Copalis" = "#F46D43", "Mocrocks" = "#9E0142")
DONOR_COL <- c("Copalis" = "#F46D43", "Mocrocks" = "#9E0142",
               "Copalis+Mocrocks (unavailable)" = "#9AA5AD")
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 3: DATA LOADING  (consolidated tidy CSVs, raw headers preserved)
# ═══════════════════════════════════════════════════════════════════════════════

```{r load}
data_dir <- here::here("02_data", "consolidated")
rd_raw   <- function(f) readr::read_csv(file.path(data_dir, f), show_col_types = FALSE,
                                        progress = FALSE)

dig_days     <- rd_raw("dig_days.csv")
cf_windows   <- rd_raw("cf_windows.csv")           # condition-specific CF bootstrap (daylight/evening)
cf_uncert    <- rd_raw("cf_uncertainty.csv")       # single-table CF bootstrap (one per beach)
pair_errors  <- rd_raw("imputation_pair_errors.csv")
season_est   <- rd_raw("season_estimates.csv")     # for validation + the u=1 historical-ratio floor

TARGETS <- c("Long Beach", "Twin Harbors")
DONORS  <- c("Copalis", "Mocrocks")
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 4: ERROR-MODEL PARAMETERS  (extracted from the CSVs)
# ═══════════════════════════════════════════════════════════════════════════════

The season total for a target beach is a sum over open beach-days. Each day is either
**counted** (its effort is a direct expansion of a beach count) or **imputed** (its
effort is borrowed from a northern donor). Four error sources enter, with very different
correlation structure:

1. **Correction factor (CF) — systematic.** Every target day, counted *and* imputed, is
   ultimately scaled by the target beach's own correction factor (an imputed day is
   `X_r · D_t/D_r`; the donor's CF cancels in the ratio, the target's survives in `X_r`).
   One table therefore moves the whole season together — it does **not** average down and
   sets a floor. Modelled as one shared draw per condition table the beach uses.
2. **Interview (digging fraction) — independent per counted day.** Averages down.
3. **Imputation — a tide-series block error on the imputed days.** Large per day
   (sd of log ≈ 0.55–0.64 for a single northern donor) but negatively autocorrelated
   within a tide series, so it averages down; the reduction is taken from the measured
   tide-series aggregation.
4. **CPUE transfer — imputed days only, with a shared season component.** The catch rate
   on an imputed day must be borrowed; this error has an irreducible season-level piece
   that does **not** average down, and it dominates *harvest* variance at LB and TH.

```{r params}
# ---- 4.1 CF relative sd, single operational table (cf_uncertainty modal rows) ----------
# NB: filter() drops NA rows; logical [ ] subsetting would keep phantom all-NA rows.
cf1 <- cf_uncert %>% filter(toupper(trimws(`used as the modal window`)) == "YES")
SIG_CF_SINGLE <- setNames(num(cf1[["relative sd"]]), cf1[["beach"]])
# e.g. Long Beach 0.056, Twin Harbors 0.051, Copalis 0.126, Mocrocks 0.095 (= v10 Table 19 floors)

# ---- 4.2 CF relative sd, condition-specific tables (cf_windows modal rows) --------------
cfc <- cf_windows %>% filter(toupper(trimws(`modal window`)) == "YES")
SIG_CF_COND <- cfc %>%
  transmute(beach, condition, rel_sd = num(`relative sd`))
# e.g. LB Daylight 0.184 / Evening 0.089; TH Daylight 0.147 / Evening 0.051

cf_cond_lookup <- function(beach, condition) {
  v <- SIG_CF_COND$rel_sd[SIG_CF_COND$beach == beach & SIG_CF_COND$condition == condition]
  if (length(v) == 0 || !is.finite(v[1])) return(unname(SIG_CF_SINGLE[beach]))
  v[1]
}

# ---- 4.3 Single-day imputation log-sd by target<-donor (pair_errors main rows) ---------
donor_all <- c("Copalis", "Mocrocks", "Copalis+Mocrocks", "Twin Harbors", "Long Beach")
pe_main <- pair_errors %>%
  filter(target %in% c(TARGETS, "Copalis", "Mocrocks"), donor %in% donor_all) %>%
  transmute(target, donor, sd1 = num(`sd of log error`))
sig_imp1 <- function(target, donor) {
  v <- pe_main$sd1[pe_main$target == target & pe_main$donor == donor]
  if (length(v) == 0 || !is.finite(v[1])) return(NA_real_)
  v[1]
}

# ---- 4.4 Season-part multiplier on the single-day imputation sd ------------------------
# From pair_errors "By season part": Fall 0.616, Spring 0.508, Winter 0.522 (pooled),
# divided by the pooled single-day sd (0.541). Fall imputation is ~14% noisier than Spring.
SEASONPART_FACTOR <- c(Fall   = 0.616 / POOLED_IMP_1DAY,
                       Winter = 0.522 / POOLED_IMP_1DAY,
                       Spring = 0.508 / POOLED_IMP_1DAY)

# ---- 4.5 Tide-series aggregation shape g(k) -------------------------------------------
# How the imputation log-sd of a k-day imputed BLOCK shrinks relative to a single day.
# From pair_errors "Tide-series SUMS": k=2 -> 0.256, k=3-4 -> 0.207, k=5+ -> 0.153, over
# the single-day 0.541. Shrinkage is faster than 1/sqrt(k) because residuals within a
# series are negatively autocorrelated (lag-1 = -0.68, v10 sec 10.1).
.g_knots_k <- c(1, 2, 3.5, 6)
# k=1 anchor must be 1 (a single day vs itself, no shrinkage); ONLY the three measured
# series-sum log-sds are normalised by the single-day sd. Putting POOLED_IMP_1DAY in the
# numerator makes the first knot 0.541/0.541 = 1 (dividing a literal 1 would give 1.848).
.g_knots_g <- c(POOLED_IMP_1DAY, 0.256, 0.207, 0.153) / POOLED_IMP_1DAY  # = 1, 0.473, 0.383, 0.283
g_agg <- function(k) {
  k <- pmax(k, 1)
  as.numeric(approx(.g_knots_k, .g_knots_g, xout = k, rule = 2)$y)
}

# ---- 4.6 CPUE transfer split (shared season component + independent day component) -----
SIG_CPUE_INDEP <- sqrt(max(0, SIG_CPUE_TRANSFER^2 - SIG_CPUE_SEASON^2))  # ~0.427

# ---- 4.7 u = 1 floor: the historical target/donor season-effort ratio CV ---------------
# When a beach is 100% imputed there is no counted reference day; the season is scaled off
# the donor by the multi-year ratio, whose season-to-season drift is a systematic floor
# that donor sampling cannot reduce (v10 sec 11.9 / Figure 21).
eff_wide <- season_est %>%
  transmute(season, beach, eff = num(`effort: corrected`)) %>%
  pivot_wider(names_from = beach, values_from = eff)
ratio_floor <- function(target, donor) {
  if (donor == "Copalis+Mocrocks") {
    d <- eff_wide[["Copalis"]] + eff_wide[["Mocrocks"]]
  } else d <- eff_wide[[donor]]
  r <- eff_wide[[target]] / d
  sd(log(r), na.rm = TRUE)   # log-sd of the season ratio ~ its CV
}

param_tbl <- tibble(
  target = rep(TARGETS, each = 3),
  donor  = rep(c("Copalis", "Mocrocks", "Copalis+Mocrocks"), 2)) %>%
  mutate(single_day_imp_sd = mapply(sig_imp1, target, donor),
         u1_ratio_floor     = mapply(ratio_floor, target, donor),
         cf_single          = SIG_CF_SINGLE[target])
param_tbl
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 5: SEASON SKELETON  (real open-day structure, used as the 2026-27 template)
# ═══════════════════════════════════════════════════════════════════════════════

```{r skeleton}
# One row per open beach-day for the target beaches in the planning season.
# light condition is empty in dig_days, so we derive it from AM/PM: AM = daylight
# (spring morning tides), PM = evening/near-dark (fall & winter). effort (workbook) is
# the day's point-estimate effort (populated on every open day, imputed or not) and is
# used as the day weight. CPUE is filled from the beach's measured days when missing.
build_skeleton <- function(beach, season) {
  d <- dig_days %>%
    filter(beach == !!beach, season == !!season, status == "Open") %>%
    transmute(
      date      = as.character(date),
      series    = `tide series`,
      part      = `season part`,
      ampm      = `AM/PM`,
      condition = ifelse(`AM/PM` == "AM", "Daylight", "Evening"),
      meanE     = num(`effort (workbook)`),
      cpue_raw  = num(`CPUE (database)`),
      monitored = monitored)
  cpue_fill <- mean(d$cpue_raw, na.rm = TRUE)
  if (!is.finite(cpue_fill)) cpue_fill <- 12
  d %>%
    mutate(cpue    = ifelse(is.finite(cpue_raw), cpue_raw, cpue_fill),
           series  = factor(series, levels = unique(series)),
           part    = factor(part, levels = c("Fall", "Winter", "Spring"))) %>%
    filter(is.finite(meanE), meanE > 0) %>%
    arrange(series)
}

skeletons <- setNames(lapply(TARGETS, build_skeleton, season = planning_season), TARGETS)

# actual monitored share in the planning season (the current operating point)
actual_counted_share <- sapply(skeletons, function(s)
  mean(s$monitored == "Yes", na.rm = TRUE))
cat("Open days in", planning_season, ":\n")
print(sapply(skeletons, nrow))
cat("Actual fully-counted (monitored==Yes) share:\n")
print(round(actual_counted_share, 3))
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 6: THE MONTE CARLO ENGINE
# ═══════════════════════════════════════════════════════════════════════════════

`sim_cv()` draws `n_rep` season totals for a target beach given which days are counted,
the donor used on imputed days, and whether the CF error is a single operational table or
condition-specific. It returns the CV of effort and of harvest, and can attribute variance
to each source by toggling one source on at a time.

```{r engine}
# choose which days to count: the m largest-effort days (v10 sec 11.3, the best rule)
choose_counted <- function(skel, m) {
  m <- max(0L, min(nrow(skel), as.integer(round(m))))
  ord <- order(skel$meanE, decreasing = TRUE)
  counted <- rep(FALSE, nrow(skel)); if (m > 0) counted[ord[seq_len(m)]] <- TRUE
  counted
}

sim_cv <- function(skel, counted, target, donor, cf_model = c("single", "condition"),
                   n_rep = 3000L,
                   use_cf = TRUE, use_int = TRUE, use_imp = TRUE,
                   use_cpue_meas = TRUE, use_cpue_transfer = TRUE) {
  cf_model <- match.arg(cf_model)
  D <- nrow(skel); R <- n_rep
  meanE <- skel$meanE; meanC <- skel$cpue
  imp <- !counted                      # imputed days
  s1  <- sig_imp1(target, donor); if (!is.finite(s1)) s1 <- POOLED_IMP_1DAY

  ## --- effort log-error matrix (D x R) ---
  logE <- matrix(0, D, R)

  # (1) correction factor: SYSTEMATIC. In single-table mode ONE draw moves the whole beach
  #     season together (one operational table; it does not average down and sets the floor).
  #     In condition mode the daylight and evening tables are INDEPENDENT systematic draws, so
  #     a season that mixes morning and evening tides partly averages across the two tables.
  if (use_cf) {
    if (cf_model == "single") {
      sd_cf <- unname(SIG_CF_SINGLE[target]); if (!is.finite(sd_cf)) sd_cf <- 0.10
      logE <- logE + matrix(rnorm(R, 0, sd_cf), D, R, byrow = TRUE)  # one draw, all days
    } else {
      for (cond in unique(skel$condition)) {
        sd_cf <- cf_cond_lookup(target, cond)
        if (!is.finite(sd_cf)) sd_cf <- unname(SIG_CF_SINGLE[target])
        draw <- rnorm(R, 0, sd_cf)                    # shared across all days of this condition
        rows <- which(skel$condition == cond)
        logE[rows, ] <- logE[rows, ] + matrix(draw, length(rows), R, byrow = TRUE)
      }
    }
  }

  # (2) interview error: independent per counted day
  if (use_int && any(counted)) {
    cr <- which(counted)
    logE[cr, ] <- logE[cr, ] + matrix(rnorm(length(cr) * R, 0, SIG_INT_DAY), length(cr), R)
  }

  # (3) imputation error: one shared draw per tide-series imputed BLOCK, sd = s1 * part * g(k)
  if (use_imp && any(imp)) {
    for (sv in levels(skel$series)) {
      rows <- which(as.character(skel$series) == sv & imp)
      k <- length(rows); if (k == 0) next
      part <- as.character(skel$part[rows[1]])
      pf <- unname(SEASONPART_FACTOR[part]); if (!is.finite(pf)) pf <- 1
      tau <- s1 * pf * g_agg(k)
      draw <- rnorm(R, 0, tau)                        # shared across the block (reproduces the series-sum sd)
      logE[rows, ] <- logE[rows, ] + matrix(draw, k, R, byrow = TRUE)
    }
  }

  Emat <- meanE * exp(logE)                           # recycles meanE down columns (D x R)
  Ehat <- colSums(Emat)

  ## --- CPUE log-error matrix for harvest ---
  logC <- matrix(0, D, R)
  if (use_cpue_meas && any(counted)) {
    cr <- which(counted)
    logC[cr, ] <- logC[cr, ] + matrix(rnorm(length(cr) * R, 0, SIG_CPUE_MEAS_DAY), length(cr), R)
  }
  if (use_cpue_transfer && any(imp)) {
    ir <- which(imp)
    theta <- rnorm(R, 0, SIG_CPUE_SEASON)             # shared season component (does NOT average down)
    logC[ir, ] <- logC[ir, ] +
      matrix(theta, length(ir), R, byrow = TRUE) +
      matrix(rnorm(length(ir) * R, 0, SIG_CPUE_INDEP), length(ir), R)
  }
  Hmat <- Emat * (meanC * exp(logC))
  Hhat <- colSums(Hmat)

  list(cv_eff  = sd(Ehat) / mean(Ehat),
       cv_harv = sd(Hhat) / mean(Hhat),
       mean_eff = mean(Ehat), mean_harv = mean(Hhat))
}
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 7: VALIDATION  (reproduce the published 2025-26 CVs and variance shares)
# ═══════════════════════════════════════════════════════════════════════════════

Before using the model to explore staffing, we check it reproduces the *published*
2025-26 season CVs (from `season_estimates.csv`) and the variance shares (from
`variance_decomposition.csv`) at each beach's **actual** counted share, using the
single operational table (the current practice) and the realistic donor.

```{r validation}
set.seed(42)
val <- lapply(TARGETS, function(b) {
  skel <- skeletons[[b]]
  m_actual <- round(actual_counted_share[b] * nrow(skel))
  counted  <- choose_counted(skel, m_actual)
  donor    <- "Copalis"   # representative single northern donor
  full <- sim_cv(skel, counted, b, donor, "single", n_rep)
  # variance shares (single source at a time; squared CVs normalised)
  src <- function(...) sim_cv(skel, counted, b, donor, "single", n_rep, ...)
  e_cf  <- src(use_cf=TRUE,  use_int=FALSE, use_imp=FALSE, use_cpue_meas=FALSE, use_cpue_transfer=FALSE)$cv_eff
  e_int <- src(use_cf=FALSE, use_int=TRUE,  use_imp=FALSE, use_cpue_meas=FALSE, use_cpue_transfer=FALSE)$cv_eff
  e_imp <- src(use_cf=FALSE, use_int=FALSE, use_imp=TRUE,  use_cpue_meas=FALSE, use_cpue_transfer=FALSE)$cv_eff
  h_cf  <- src(use_cf=TRUE,  use_int=FALSE, use_imp=FALSE, use_cpue_meas=FALSE, use_cpue_transfer=FALSE)$cv_harv
  h_int <- src(use_cf=FALSE, use_int=TRUE,  use_imp=FALSE, use_cpue_meas=TRUE,  use_cpue_transfer=FALSE)$cv_harv
  h_imp <- src(use_cf=FALSE, use_int=FALSE, use_imp=TRUE,  use_cpue_meas=FALSE, use_cpue_transfer=FALSE)$cv_harv
  h_cpt <- src(use_cf=FALSE, use_int=FALSE, use_imp=FALSE, use_cpue_meas=FALSE, use_cpue_transfer=TRUE )$cv_harv
  esum <- e_cf^2 + e_int^2 + e_imp^2
  hsum <- h_cf^2 + h_int^2 + h_imp^2 + h_cpt^2
  pub <- season_est %>% filter(season == planning_season, beach == b)
  tibble(beach = b,
         counted_share = round(m_actual / nrow(skel), 2),
         cv_eff_model = round(full$cv_eff, 3),
         cv_eff_pub   = round(num(pub[["effort CV"]]), 3),
         cv_harv_model = round(full$cv_harv, 3),
         cv_harv_pub   = round(num(pub[["harvest CV"]]), 3),
         eff_cf = round(e_cf^2/esum, 2), eff_imp = round(e_imp^2/esum, 2),
         harv_cf = round(h_cf^2/hsum, 2), harv_imp = round(h_imp^2/hsum, 2),
         harv_cpue_transfer = round(h_cpt^2/hsum, 2))
}) %>% bind_rows()
val
```

The model should land close to the published effort CVs (LB ~0.07, TH ~0.06) and harvest
CVs (LB ~0.14, TH ~0.10), and reproduce the qualitative variance split reported in the
published `variance_decomposition.csv` (which is cited here for comparison, not read by this
document): 2025-26 LB effort ~49% CF / ~49% imputation; LB harvest dominated by CPUE
transfer ~53%; TH harvest CPUE transfer ~63%. Differences reflect that
this is an independent Monte Carlo reconstruction, not the source workbook's own bootstrap;
treat absolute levels as approximate and the **relative** staffing comparisons as the
product.

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 8: FIGURE 1 — CV vs how much the beach is imputed, by donor
# ═══════════════════════════════════════════════════════════════════════════════

**Two distinct regimes — do not confuse them.** The swept curves below are *within-season*
imputation: on an uncounted day the workbook borrows a **nearby counted reference day's**
day-over-day ratio, an error that is large per day but negatively autocorrelated within a
tide series, so it averages down (this is why counting even a few of the biggest days keeps
effort CV low). That is a different thing from *full* imputation (`u = 1`, a budget
elimination), where the beach is counted on **no** day, there is no reference day, and the
whole season is scaled off the donor by the multi-year **average** ratio — whose
season-to-season drift is a systematic floor that does not average down (v10 §11.9 /
Figure 21). On the effort panel this `u = 1` floor is drawn as a **star**; the swept curve
must not be read all the way to a 100% cost, because the mechanism changes there.

```{r sweep}
# sweep the number of counted days from a floor of 1 up to fully counted, for each
# target x donor, single operational table. Also record the u=1 ratio floor.
donor_lab <- c("Copalis" = "Copalis", "Mocrocks" = "Mocrocks",
               "Copalis+Mocrocks" = "Copalis+Mocrocks (unavailable)")

sweep_one <- function(target, donor, cf_model = "single") {
  skel <- skeletons[[target]]; N <- nrow(skel)
  ms <- unique(round(seq(1, N, length.out = n_grid)))
  purrr::map_dfr(ms, function(m) {
    counted <- choose_counted(skel, m)
    r <- sim_cv(skel, counted, target, donor, cf_model, n_rep)
    imp_eff_share <- sum(skel$meanE[!counted]) / sum(skel$meanE)
    tibble(target, donor = unname(donor_lab[donor]), cf_model,
           m_counted = m, pct_counted = m / N,
           imputed_share = 1 - m / N, imputed_eff_share = imp_eff_share,
           cv_eff = r$cv_eff, cv_harv = r$cv_harv)
  })
}

set.seed(42)
grid <- purrr::map_dfr(TARGETS, function(t)
  purrr::map_dfr(c("Copalis", "Mocrocks", "Copalis+Mocrocks"),
                 ~ sweep_one(t, .x, "single")))

grid_long <- grid %>%
  pivot_longer(c(cv_eff, cv_harv), names_to = "metric", values_to = "cv") %>%
  mutate(metric = recode(metric, cv_eff = "Effort", cv_harv = "Harvest"))

# u = 1 endpoint: the FULL-imputation (budget-elimination) floor — a systematic donor-ratio
# drift, a different regime from the within-season sweep. Marked only on the effort panel.
full_imp <- param_tbl %>%
  transmute(target, donor = unname(donor_lab[donor]), metric = "Effort",
            imputed_share = 1, cv = u1_ratio_floor)

fig1 <- ggplot(grid_long,
               aes(imputed_share, cv, color = donor, linetype = donor)) +
  geom_hline(yintercept = CV_DESIGN, linetype = "dashed", color = "#4a8f5b") +
  geom_hline(yintercept = CV_MRIP,   linetype = "dashed", color = "#C1666B") +
  geom_line(linewidth = 1) + geom_point(size = 1.4) +
  geom_point(data = full_imp, aes(imputed_share, cv, color = donor),
             shape = 8, size = 3, stroke = 1.1, inherit.aes = FALSE, show.legend = FALSE) +
  facet_grid(metric ~ target, scales = "free_y") +
  scale_x_continuous("Share of the beach's open days that are IMPUTED (borrowed from one donor)",
                     labels = percent_format(accuracy = 1)) +
  scale_y_continuous("Season estimate CV") +
  scale_color_manual(values = DONOR_COL, name = "Single donor") +
  scale_linetype_manual(values = c("Copalis" = "solid", "Mocrocks" = "solid",
                                   "Copalis+Mocrocks (unavailable)" = "dotted"),
                        name = "Single donor") +
  labs(title = "Season CV rises with the imputed share — and the donor choice matters",
       subtitle = paste0("Long Beach & Twin Harbors, ", planning_season,
                         " season shape. Green = 0.20 design target, red = MRIP 0.30. Joint donor ",
                         "(grey, dotted) is unbuildable under 2-on/2-off co-management. Stars on the ",
                         "effort row = the u=1 full-imputation (budget-cut) floor, a separate regime.")) +
  my_theme
fig1
save_plot(fig1, "cv_curves", "fig1_cv_vs_imputed_by_donor.png", w = 11, h = 8)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 9: FIGURE 2 — does modelling daylight vs evening differently change the answer?
# ═══════════════════════════════════════════════════════════════════════════════

The current practice applies **one** correction-factor table per beach. The survey record
supports separate **daylight** and **evening** tables. For LB and TH the daylight tables
rest on only 3 and 5 old surveys, so splitting by condition *raises* their CF error unless
new egress surveys are run. This figure contrasts the two CF models on the effort CV.

```{r cf_model}
set.seed(42)
grid_cf <- purrr::map_dfr(TARGETS, function(t)
  purrr::map_dfr(c("single", "condition"), ~ sweep_one(t, "Copalis", .x))) %>%
  mutate(cf_lab = recode(cf_model, single = "One operational table (current)",
                         condition = "Separate daylight / evening tables"))

fig2 <- ggplot(grid_cf, aes(imputed_share, cv_eff, color = cf_lab)) +
  geom_hline(yintercept = CV_DESIGN, linetype = "dashed", color = "#4a8f5b") +
  geom_line(linewidth = 1) + geom_point(size = 1.3) +
  facet_wrap(~ target) +
  scale_x_continuous("Imputed share of open days", labels = percent_format(accuracy = 1)) +
  scale_y_continuous("Season EFFORT CV") +
  scale_color_manual(values = c("One operational table (current)" = "#1B4965",
                                "Separate daylight / evening tables" = "#E0A458"),
                     name = NULL) +
  labs(title = "Splitting the correction factor by light condition exposes thin daylight tables",
       subtitle = "Long Beach & Twin Harbors have only 3 and 5 daylight surveys (newest 2011); condition-specific tables raise their CF floor until resurveyed.") +
  my_theme
fig2
save_plot(fig2, "cf_model", "fig2_cf_single_vs_condition.png", w = 11, h = 5)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 10: FIGURE 3 — season-part stratification (fall / winter / spring)
# ═══════════════════════════════════════════════════════════════════════════════

Fall openers fall on evening tides (noisier imputation, sd 0.616) and use the evening CF
table; spring openers are morning/daylight (imputation sd 0.508) and use the daylight
table. We re-run the sweep on each season-part sub-skeleton to show where imputation is
most affordable.

```{r season_part}
sweep_part <- function(target, part, donor = "Copalis", cf_model = "single") {
  full <- skeletons[[target]]
  skel <- full %>% filter(part == !!part)
  if (nrow(skel) < 2) return(NULL)
  N <- nrow(skel); ms <- unique(round(seq(1, N, length.out = min(n_grid, N))))
  purrr::map_dfr(ms, function(m) {
    counted <- choose_counted(skel, m)
    r <- sim_cv(skel, counted, target, donor, cf_model, n_rep)
    tibble(target, part, imputed_share = 1 - m / N,
           cv_eff = r$cv_eff, cv_harv = r$cv_harv)
  })
}

set.seed(42)
grid_part <- purrr::map_dfr(TARGETS, function(t)
  purrr::map_dfr(c("Fall", "Winter", "Spring"), ~ sweep_part(t, .x)))

fig3 <- ggplot(grid_part, aes(imputed_share, cv_harv, color = part)) +
  geom_hline(yintercept = CV_DESIGN, linetype = "dashed", color = "#4a8f5b") +
  geom_line(linewidth = 1) + geom_point(size = 1.3) +
  facet_wrap(~ target) +
  scale_x_continuous("Imputed share of that part's open days", labels = percent_format(accuracy = 1)) +
  scale_y_continuous("Season-part HARVEST CV") +
  scale_color_manual(values = c(Fall = "#C1666B", Winter = "#1B4965", Spring = "#4a8f5b"),
                     name = "Season part") +
  labs(title = "Which part of the season tolerates imputation best?",
       subtitle = "Fall (evening tides) carries the noisiest imputation; spring (morning/daylight) the least. Donor = Copalis, one operational table.") +
  my_theme
fig3
save_plot(fig3, "season_part", "fig3_cv_by_season_part.png", w = 11, h = 5)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 11: FIGURE 4 + TABLE — maximum imputable share to hit each CV target
# ═══════════════════════════════════════════════════════════════════════════════

Inverting the curves: for each target x donor x metric, the largest imputed share that
still meets a CV target. This is the staffing-decision table — "how few days can we
count and still publish?"

```{r decision}
max_imputable <- function(sub, cv_target) {
  # sub: rows for one target/donor/metric ordered by imputed_share ascending
  ok <- sub$cv <= cv_target
  if (!any(ok)) return(0)
  max(sub$imputed_share[ok])
}

dec <- grid_long %>%
  filter(donor != "Copalis+Mocrocks (unavailable)") %>%
  group_by(target, donor, metric) %>%
  arrange(imputed_share, .by_group = TRUE) %>%
  group_modify(~ tibble(
    `CV<=0.15` = max_imputable(.x, CV_TIGHT),
    `CV<=0.20` = max_imputable(.x, CV_DESIGN),
    `CV<=0.30` = max_imputable(.x, CV_MRIP))) %>%
  ungroup()

dec_long <- dec %>%
  pivot_longer(starts_with("CV<="), names_to = "cv_target", values_to = "max_imp")

fig4 <- ggplot(dec_long,
               aes(cv_target, interaction(metric, donor, sep = " · "), fill = max_imp)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = percent(max_imp, accuracy = 1)), size = 4) +
  facet_wrap(~ target) +
  scale_fill_gradient2("Max imputable share", low = "#C1666B", mid = "#E0A458",
                       high = "#4a8f5b", midpoint = 0.5, labels = percent) +
  labs(title = "How much of each beach may be imputed and still meet a CV target",
       subtitle = "Higher (green) = more days can be borrowed. Effort tolerates far more imputation than harvest, because harvest carries the un-averaging CPUE-transfer error.",
       x = "CV target", y = NULL) +
  my_theme + theme(legend.position = "right")
fig4
save_plot(fig4, "decision", "fig4_max_imputable_share.png", w = 12, h = 6)

readr::write_csv(dec_long, file.path(folder_name, "tables", "decision_max_imputable_share.csv"))
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 12: FIGURE 5 — crew-day frontier for the 2026-27 plan
# ═══════════════════════════════════════════════════════════════════════════════

Translate to field cost. A Long Beach full count is two events (a north and a south
segment walk), so it costs **2 crew-day units**; a Twin Harbors count costs **1**.
Copalis and Mocrocks are held at their co-management floor (one of the pair counted every
state dig day) and are not the lever here. We hold harvest CV as the reported metric and
trace the achievable CV against the crew-day budget spent counting LB and TH.

```{r crew}
crew_unit <- c("Long Beach" = 2, "Twin Harbors" = 1)

crew_curve <- function(target, donor = "Copalis") {
  skel <- skeletons[[target]]; N <- nrow(skel); u <- crew_unit[target]
  ms <- unique(round(seq(1, N, length.out = n_grid)))
  purrr::map_dfr(ms, function(m) {
    counted <- choose_counted(skel, m)
    r <- sim_cv(skel, counted, target, donor, "single", n_rep)
    tibble(target, m_counted = m, crew_days = m * u,
           cv_eff = r$cv_eff, cv_harv = r$cv_harv)
  })
}

set.seed(42)
crew <- purrr::map_dfr(TARGETS, crew_curve)
status_quo <- tibble(
  target = TARGETS,
  crew_days = sapply(TARGETS, function(b)
    round(actual_counted_share[b] * nrow(skeletons[[b]])) * crew_unit[b]))

fig5 <- ggplot(crew, aes(crew_days, cv_harv, color = target)) +
  geom_hline(yintercept = CV_DESIGN, linetype = "dashed", color = "#4a8f5b") +
  geom_hline(yintercept = CV_MRIP,   linetype = "dashed", color = "#C1666B") +
  geom_line(linewidth = 1) + geom_point(size = 1.3) +
  geom_vline(data = status_quo, aes(xintercept = crew_days, color = target),
             linetype = "dotted", linewidth = 0.8, show.legend = FALSE) +
  scale_color_manual(values = BEACH_COL, name = "Target beach") +
  labs(title = "Crew-day frontier: harvest CV vs days spent counting (2026-27 shape)",
       subtitle = "LB counted day = 2 crew-units (two-section walk); TH = 1. Dotted verticals mark the 2025-26 status quo. Donor = Copalis, one operational table.",
       x = "Crew-day units spent counting the target beach",
       y = "Season HARVEST CV") +
  my_theme
fig5
save_plot(fig5, "crew_days", "fig5_crew_day_frontier.png", w = 11, h = 6)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 13: FIGURE 6 — variance composition as imputation rises (Long Beach harvest)
# ═══════════════════════════════════════════════════════════════════════════════

Why harvest is hard to buy down: as the imputed share rises, the CPUE-transfer term (with
its irreducible season component) takes over the harvest variance, so counting a few more
days barely moves it — the lever is measured CPUE and egress surveys, not day-counts.

```{r variance_comp}
set.seed(42)
comp <- function(target, donor = "Copalis") {
  skel <- skeletons[[target]]; N <- nrow(skel)
  ms <- unique(round(seq(1, N, length.out = n_grid)))
  purrr::map_dfr(ms, function(m) {
    counted <- choose_counted(skel, m)
    s <- function(...) sim_cv(skel, counted, target, donor, "single", n_rep, ...)$cv_harv
    cf  <- s(use_cf=TRUE,  use_int=FALSE, use_imp=FALSE, use_cpue_meas=FALSE, use_cpue_transfer=FALSE)
    int <- s(use_cf=FALSE, use_int=TRUE,  use_imp=FALSE, use_cpue_meas=TRUE,  use_cpue_transfer=FALSE)
    im  <- s(use_cf=FALSE, use_int=FALSE, use_imp=TRUE,  use_cpue_meas=FALSE, use_cpue_transfer=FALSE)
    cpt <- s(use_cf=FALSE, use_int=FALSE, use_imp=FALSE, use_cpue_meas=FALSE, use_cpue_transfer=TRUE )
    tot <- cf^2 + int^2 + im^2 + cpt^2
    tibble(target, imputed_share = 1 - m / N,
           `Correction factor` = cf^2/tot, `Interviews / measured CPUE` = int^2/tot,
           `Imputation` = im^2/tot, `CPUE transfer` = cpt^2/tot)
  })
}
comp_lb <- comp("Long Beach") %>%
  pivot_longer(-c(target, imputed_share), names_to = "source", values_to = "share")

fig6 <- ggplot(comp_lb, aes(imputed_share, share, fill = source)) +
  geom_area(color = "white", linewidth = 0.2) +
  scale_x_continuous("Imputed share of open days", labels = percent_format(accuracy = 1)) +
  scale_y_continuous("Share of season HARVEST variance", labels = percent) +
  scale_fill_manual(values = c("Correction factor" = "#1B4965",
                               "Interviews / measured CPUE" = "#4A7BA6",
                               "Imputation" = "#E0A458", "CPUE transfer" = "#C1666B"),
                    name = NULL) +
  labs(title = "Long Beach harvest: CPUE transfer takes over as imputation rises",
       subtitle = "The CPUE-transfer error has a season-level component that does not average down — so counting more days cannot remove it; measured CPUE and egress surveys can.") +
  my_theme + theme(legend.position = "right")
fig6
save_plot(fig6, "variance", "fig6_variance_composition_lb.png", w = 11, h = 6)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 14: WRITE THE FULL CV GRID + PARAMETERS
# ═══════════════════════════════════════════════════════════════════════════════

```{r write_outputs}
readr::write_csv(grid,       file.path(folder_name, "tables", "cv_grid_by_donor.csv"))
readr::write_csv(grid_cf,    file.path(folder_name, "tables", "cv_grid_cf_model.csv"))
readr::write_csv(grid_part,  file.path(folder_name, "tables", "cv_grid_season_part.csv"))
readr::write_csv(crew,       file.path(folder_name, "tables", "crew_day_frontier.csv"))
readr::write_csv(param_tbl,  file.path(folder_name, "tables", "model_parameters.csv"))
readr::write_csv(val,        file.path(folder_name, "validation", "validation_vs_published.csv"))
cat("Wrote CV grids, decision table, parameters, and validation to", folder_name, "\n")
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 15: HEADLINE NUMBERS FOR THE 2026-27 STAFFING DECISION
# ═══════════════════════════════════════════════════════════════════════════════

```{r headline}
# For each target/donor: the counted share needed to hit the 0.20 harvest design target,
# and the donor penalty vs the (unavailable) joint donor.
answer <- grid %>%
  group_by(target, donor) %>%
  arrange(imputed_share, .by_group = TRUE) %>%
  summarise(
    min_counted_for_harvest_0.20 = {
      ok <- cv_harv <= CV_DESIGN
      if (any(ok)) 1 - max(imputed_share[ok]) else NA_real_
    },
    min_counted_for_effort_0.20 = {
      ok <- cv_eff <= CV_DESIGN
      if (any(ok)) 1 - max(imputed_share[ok]) else NA_real_
    }, .groups = "drop") %>%
  mutate(across(starts_with("min_counted"), ~ round(.x, 2)))
answer

cat("\n--- 2026-27 staffing read-out (", planning_season, "skeleton ) ---\n")
cat("Actual", planning_season, "counted share:  Long Beach",
    round(actual_counted_share["Long Beach"], 2),
    "| Twin Harbors", round(actual_counted_share["Twin Harbors"], 2), "\n")
cat("u=1 (full-imputation) systematic floors, log-sd of the season ratio:\n")
print(param_tbl %>% select(target, donor, u1_ratio_floor) %>%
        mutate(u1_ratio_floor = round(u1_ratio_floor, 3)))
cat("\nInterpretation: for EFFORT the CF floor and averaging-down imputation let a beach be\n",
    "heavily borrowed and still meet 0.20; for HARVEST the CPUE-transfer floor binds, so\n",
    "harvest is the constraint that sets how few days may be counted. Copalis is the better\n",
    "single donor for Twin Harbors; the two donors are close for Long Beach. The joint donor\n",
    "would beat both but cannot be built under co-management.\n")
```
