---
title: "Razor Clam CPUE Cross-Beach Transfer — how well does borrowing catch rate from one beach match another? (2026-27 planning)"
author: "WDFW Coastal Shellfish Unit"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    df_print: paged
---

<!--
  PURPOSE
  On a day a beach is not counted, the workbook imputes its EFFORT from a northern donor
  (Copalis / Mocrocks) and, separately, must borrow a CATCH RATE (CPUE) to turn that effort
  into harvest. CPUE transfer is the single largest source of HARVEST variance at Long Beach
  and Twin Harbors (v10 sec 10.1). This document measures, directly from the record, how well
  a borrowed catch rate actually matches the target beach, and asks whether a dedicated CPUE
  study is warranted. It answers four questions the v10 document only summarises with one
  number (0.448):

    (1) BY DONOR: which beach's catch rate best matches Long Beach / Twin Harbors, and is the
        transfer biased?
    (2) BY LAG: is CPUE transfer a same-day phenomenon, or does a nearby day do just as well?
    (3) SEASON-LEVEL: how much of the transfer error is a systematic season-level piece that
        does not average down (the v10 model's 0.137)?
    (4) OWN vs BORROW: how stable is a beach's OWN catch rate across the days of a tide series,
        i.e. how much would interviewing the target beach itself buy over borrowing a donor?

  STRUCTURE follows 20260710-razor-clam-harvest-figures-pipeline.Rmd and the co-management
  staffing model: load.lib loop, my_theme, dated 03_analyses/ output tree, save_plot dpi=300.
  Reads ONLY the tidy CSVs in 02_data/consolidated/. Measured CPUE = a beach-day with at least
  one creel interview (n_interviews > 0) and CPUE > 0 in harvest_by_beach_day.csv; hand-typed
  CPUE on estimated days is excluded, so every ratio below is measured-vs-measured.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r reset}
rm(list = ls())
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 0: TOGGLES
# ═══════════════════════════════════════════════════════════════════════════════

```{r toggles}
save_fig  <- TRUE
out_date  <- "20260714"     # pin the output-folder date for a reproducible patch (= Sys.Date())
MIN_SEASON_N <- 2L          # min measured pairs in a season to use it in the season-drift estimate
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 1: PACKAGES, THEME, HELPERS
# ═══════════════════════════════════════════════════════════════════════════════

```{r packages}
load.lib <- c("tidyverse", "readr", "scales", "here", "patchwork")
install.lib <- load.lib[!load.lib %in% installed.packages()]
for (lib in install.lib) install.packages(lib, dependencies = TRUE)
sapply(load.lib, require, character.only = TRUE)

my_theme <- theme(text             = element_text(size = 14, color = "black"),
                  panel.background = element_blank(),
                  panel.grid.major = element_line(color = "grey92", linewidth = 0.4),
                  panel.grid.minor = element_blank(),
                  axis.text        = element_text(size = 13, color = "black"),
                  panel.border     = element_rect(color = "black", fill = NA, linewidth = 1.0),
                  strip.background = element_rect(fill = "grey95", color = "black"),
                  strip.text       = element_text(size = 13, face = "bold"),
                  legend.key       = element_blank(),
                  legend.position  = "bottom",
                  plot.title       = element_text(face = "bold"),
                  plot.margin      = margin(10, 20, 10, 10))
num <- function(x) suppressWarnings(as.numeric(x))
NAVY<-"#1B4965"; RED<-"#C1666B"; GREEN<-"#4a8f5b"; GREY<-"#9AA5AD"; INK<-"#20303A"
COPc<-"#F46D43"; MOCc<-"#9E0142"
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 2: OUTPUT FOLDER
# ═══════════════════════════════════════════════════════════════════════════════

```{r out_folder}
folder_name <- here::here("03_analyses", paste0(out_date, "-cpue-cross-beach"))
for (sf in c("transfer", "lag", "own_vs_borrow", "season_drift", "tables"))
  dir.create(file.path(folder_name, sf), recursive = TRUE, showWarnings = FALSE)
save_plot <- function(p, sub, name, w = 12, h = 6.2) {
  if (save_fig) ggsave(file.path(folder_name, sub, name), plot = p, width = w, height = h, dpi = 300)
}
cat("Output folder:", folder_name, "\n")
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 3: DATA — measured CPUE, one row per (beach, date)
# ═══════════════════════════════════════════════════════════════════════════════

```{r load}
data_dir <- here::here("02_data", "consolidated")
hbd <- readr::read_csv(file.path(data_dir, "harvest_by_beach_day.csv"), show_col_types = FALSE) %>%
  mutate(CPUE = num(CPUE), n_interviews = num(n_interviews), date = as.Date(date)) %>%
  filter(!is.na(CPUE), CPUE > 0, !is.na(n_interviews), n_interviews > 0) %>%
  distinct(beach, date, .keep_all = TRUE)

wide      <- hbd %>% select(beach, date, CPUE) %>% pivot_wider(names_from = beach, values_from = CPUE)
seasonmap <- hbd %>% distinct(date, season)
cat("Measured-CPUE beach-days:", nrow(hbd), "over", n_distinct(hbd$season), "seasons\n")
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 4: SAME-DAY TRANSFER ERROR BY TARGET <- DONOR  (Fig 1 + table)
# ═══════════════════════════════════════════════════════════════════════════════

The transfer error is the standard deviation of `log(donor CPUE / target CPUE)` on days both
beaches were interviewed. The median ratio measures bias.

```{r pairs}
pair_stat <- function(tgt, dnr) {
  if (dnr == "Copalis+Mocrocks") {
    d <- wide %>% filter(!is.na(.data[[tgt]]), !is.na(Copalis), !is.na(Mocrocks)) %>%
      mutate(lr = log(((Copalis + Mocrocks) / 2) / .data[[tgt]]))
  } else {
    d <- wide %>% filter(!is.na(.data[[tgt]]), !is.na(.data[[dnr]])) %>%
      mutate(lr = log(.data[[dnr]] / .data[[tgt]]))
  }
  tibble(target = tgt, donor = dnr, n = nrow(d), sd_log = sd(d$lr),
         median_ratio = median(exp(d$lr)), mape = 100 * mean(abs(exp(d$lr) - 1)))
}
pairs <- tribble(~t, ~d,
  "Long Beach","Copalis", "Long Beach","Mocrocks", "Long Beach","Twin Harbors", "Long Beach","Copalis+Mocrocks",
  "Twin Harbors","Copalis", "Twin Harbors","Mocrocks", "Twin Harbors","Long Beach", "Twin Harbors","Copalis+Mocrocks",
  "Copalis","Mocrocks")
ps <- purrr::map2_dfr(pairs$t, pairs$d, pair_stat)
ps

f1 <- ps %>% filter(target %in% c("Long Beach","Twin Harbors")) %>%
  mutate(donlab = recode(donor, "Copalis+Mocrocks"="Copalis+Mocrocks\n(joint, unavailable)",
                         "Twin Harbors"="Twin Harbors\n(southern sibling)",
                         "Long Beach"="Long Beach\n(southern sibling)"),
         region = case_when(grepl("joint", donlab) ~ "joint (unavailable)",
                            donor %in% c("Copalis","Mocrocks") ~ "northern donor (cross-divide)",
                            TRUE ~ "southern sibling"))
fig1 <- ggplot(f1, aes(reorder(donlab, sd_log), sd_log, fill = region)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = sprintf("%.2f", sd_log)), vjust = -0.4, size = 3.6, fontface = "bold", color = INK) +
  geom_hline(yintercept = ps$sd_log[ps$target=="Copalis" & ps$donor=="Mocrocks"], linetype = "dashed", color = COPc) +
  annotate("text", x = 0.7, y = 0.285, hjust = 0, size = 3, color = COPc,
           label = "Copalis and Mocrocks agree at 0.26 (same region)") +
  facet_wrap(~ target, scales = "free_x") +
  scale_fill_manual(values = c("northern donor (cross-divide)"=RED, "southern sibling"=GREEN, "joint (unavailable)"=GREY), name = NULL) +
  scale_y_continuous(limits = c(0, 0.56), expand = expansion(mult = c(0, 0.02))) +
  labs(x = NULL, y = "CPUE transfer error (sd of log donor/target)",
       title = "Borrowing catch rate across the north-south divide is the worst case",
       subtitle = "Same-day CPUE transfer error by donor; all unbiased (median ratio ~1.0). The two northern beaches match each other (0.26) far better than the southern targets.") +
  my_theme + theme(panel.grid.major.x = element_blank(), axis.text.x = element_text(size = 8.5, lineheight = 0.9),
                   plot.subtitle = element_text(size = 9.5, color = "grey30"))
fig1
save_plot(fig1, "transfer", "cpue_fig1_by_pair.png", w = 12, h = 6.2)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 5: TRANSFER ERROR vs LAG  (Fig 2)
# ═══════════════════════════════════════════════════════════════════════════════

```{r lag}
lagstat <- function(tgt, dnr, L) {
  td <- wide %>% filter(!is.na(.data[[tgt]])) %>% transmute(date, tc = .data[[tgt]])
  dd <- wide %>% filter(!is.na(.data[[dnr]])) %>% transmute(date, dc = .data[[dnr]])
  lr <- purrr::map_dbl(seq_len(nrow(td)), function(i) {
    offs <- if (L == 0) 0 else c(-L, L)
    hit <- dd$dc[dd$date %in% (td$date[i] + offs)]
    if (length(hit) == 0) return(NA_real_)
    log(hit[1] / td$tc[i])
  })
  lr <- lr[is.finite(lr)]
  if (length(lr) < 3) return(NA_real_)
  sd(lr)
}
lag_grid <- expand_grid(
    pair = c("Long Beach<-Copalis","Long Beach<-Mocrocks","Twin Harbors<-Copalis","Twin Harbors<-Mocrocks","Copalis<-Mocrocks"),
    lag = 0:3) %>%
  separate(pair, c("target","donor"), sep = "<-", remove = FALSE) %>%
  mutate(sd_log = purrr::pmap_dbl(list(target, donor, lag), lagstat))

fig2 <- ggplot(lag_grid, aes(lag, sd_log, color = pair, group = pair)) +
  geom_line(linewidth = 1) + geom_point(size = 2.4) +
  scale_color_manual(values = c("Long Beach<-Copalis"=COPc, "Long Beach<-Mocrocks"=MOCc,
                                "Twin Harbors<-Copalis"="#F4A582", "Twin Harbors<-Mocrocks"="#B2182B",
                                "Copalis<-Mocrocks"=GREY), name = NULL) +
  scale_x_continuous(breaks = 0:3) + scale_y_continuous(limits = c(0, 0.86)) +
  labs(x = "Lag between the target day and the borrowed donor CPUE (days)", y = "CPUE transfer error (sd of log)",
       title = "CPUE transfer is a same-day phenomenon; it decays within one day",
       subtitle = "Catch rate is driven by the day (tide, weather, how the dig went) more than by the beach: two beaches on one tide agree better than one beach on two tides.") +
  my_theme + theme(plot.subtitle = element_text(size = 9.5, color = "grey30"))
fig2
save_plot(fig2, "lag", "cpue_fig2_by_lag.png", w = 12, h = 6.2)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 6: SEASON-LEVEL DRIFT  (Fig 4 + table)
# ═══════════════════════════════════════════════════════════════════════════════

```{r drift}
drift <- function(tgt, dnr) {
  wide %>% filter(!is.na(.data[[tgt]]), !is.na(.data[[dnr]])) %>%
    left_join(seasonmap, by = "date") %>%
    mutate(lr = log(.data[[dnr]] / .data[[tgt]])) %>%
    group_by(season) %>% summarise(n = n(), mlr = mean(lr), .groups = "drop") %>%
    filter(n >= MIN_SEASON_N) %>% mutate(pair = paste0(tgt, " <- ", dnr))
}
dr  <- bind_rows(drift("Long Beach","Copalis"), drift("Long Beach","Mocrocks"),
                 drift("Twin Harbors","Copalis"), drift("Twin Harbors","Mocrocks"))
drsd <- dr %>% group_by(pair) %>% summarise(season_level_sd = sd(mlr), .groups = "drop")
drsd

fig4 <- ggplot(dr, aes(season, exp(mlr), group = pair, color = pair)) +
  geom_hline(yintercept = 1, linetype = "dotted", color = GREY) +
  geom_line(linewidth = 0.9) + geom_point(size = 1.8) +
  scale_color_manual(values = c("Long Beach <- Copalis"=COPc, "Long Beach <- Mocrocks"=MOCc,
                                "Twin Harbors <- Copalis"="#F4A582", "Twin Harbors <- Mocrocks"="#B2182B"), name = NULL) +
  scale_y_continuous("Season-mean donor / target CPUE ratio") +
  labs(x = NULL, title = "A season-level piece of the CPUE gap does not average down",
       subtitle = "Season-mean catch-rate ratio by pair; its drift (log-sd ~0.13-0.18) is the irreducible season component, matching the v10 model's 0.137 for Twin Harbors but higher for Long Beach.") +
  my_theme + theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8),
                   plot.subtitle = element_text(size = 9.3, color = "grey30"))
fig4
save_plot(fig4, "season_drift", "cpue_fig4_season_drift.png", w = 12, h = 6.2)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 7: WHAT REMOVES THE HARVEST FLOOR — BORROW vs INTERVIEW THE TARGET (Fig 3)
# ═══════════════════════════════════════════════════════════════════════════════

Catch rate is volatile day to day (a beach's own CPUE changes about as much from one tide to
the next as it differs from a donor's on the same tide). That day-level noise, whether ~0.45
borrowed or comparable own, is idiosyncratic and **averages down** over a season. What does
**not** average down is the season-level piece measured in Section 6 (~0.13 for Twin Harbors,
~0.16-0.18 for Long Beach). That season-level component is the harvest floor. Borrowing from a
donor cannot remove it, because the donor-to-target ratio itself drifts. Interviewing the target
beach removes it, because you measure the target's own catch rate instead of inferring it.

```{r floor}
# borrow floor: the season-level CPUE-transfer sd (mean over the four LB/TH<-northern pairs)
borrow_floor <- mean(drsd$season_level_sd)
# interview residual: with the target interviewed across the season, the transfer term vanishes;
# what remains is the sampling error of the measured season CPUE. The per-day measured-CPUE sd is
# ~0.17 (v10 sec 10.1); interviewing on ~15 spread days per season leaves a season-mean sampling
# sd of ~0.17/sqrt(15) ~ 0.04. Shown as an order-of-magnitude residual, not a claim of precision.
interview_resid <- 0.17 / sqrt(15)

f3 <- tibble(
  option = factor(c("Borrow a donor's catch rate\n(cross-beach transfer)",
                    "Interview the target beach\n(measure it directly)"),
                  levels = c("Borrow a donor's catch rate\n(cross-beach transfer)",
                             "Interview the target beach\n(measure it directly)")),
  sd_log = c(round(borrow_floor, 3), round(interview_resid, 3)),
  kind   = c("borrow", "own"))
fig3 <- ggplot(f3, aes(option, sd_log, fill = kind)) +
  geom_col(width = 0.55) +
  geom_text(aes(label = sprintf("%.2f", sd_log)), vjust = -0.4, size = 5.4, fontface = "bold", color = INK) +
  scale_fill_manual(values = c(own = GREEN, borrow = RED), guide = "none") +
  scale_y_continuous("Season-level CPUE error in harvest",
                     limits = c(0, 0.21), expand = expansion(mult = c(0, 0.03))) +
  labs(x = NULL,
       title = "Only interviewing the target beach removes the harvest floor",
       subtitle = "Day-to-day catch-rate noise averages down either way; the season-level piece (~0.15) does not, and only a direct interview removes it.") +
  my_theme + theme(panel.grid.major.x = element_blank(),
                   axis.text.x = element_text(face = "bold", size = 11, lineheight = 0.95),
                   plot.subtitle = element_text(size = 9.8, color = "grey30"))
fig3
save_plot(fig3, "own_vs_borrow", "cpue_fig3_floor.png", w = 11, h = 6.0)
cat(sprintf("harvest CPUE floor: borrow ~%.3f (season-level drift) vs interview-the-target ~%.3f (sampling residual)\n",
            borrow_floor, interview_resid))
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 8: WRITE TABLES + HEADLINE
# ═══════════════════════════════════════════════════════════════════════════════

```{r write}
readr::write_csv(ps,       file.path(folder_name, "tables", "cpue_transfer_by_pair.csv"))
readr::write_csv(lag_grid, file.path(folder_name, "tables", "cpue_transfer_by_lag.csv"))
readr::write_csv(drsd,     file.path(folder_name, "tables", "cpue_season_level_sd.csv"))
readr::write_csv(f3,       file.path(folder_name, "tables", "cpue_own_vs_borrow.csv"))
readr::write_csv(
  tibble(quantity = c("day-level transfer, same day (representative)", "day-level transfer, 1 day off",
                      "season-level transfer floor (borrow, mean of 4 pairs)", "season-level residual (interview target)"),
         sd_log = c(round(mean(lag_grid$sd_log[lag_grid$lag==0 & lag_grid$pair!="Copalis<-Mocrocks"]),3),
                    round(mean(lag_grid$sd_log[lag_grid$lag==1 & lag_grid$pair!="Copalis<-Mocrocks"]),3),
                    round(borrow_floor,3), round(interview_resid,3))),
  file.path(folder_name, "tables", "cpue_ground_truth_summary.csv"))

cat("\n=== CPUE cross-beach transfer: headline ===\n")
cat("* Day-level transfer error (LB/TH from a northern donor): ~0.43-0.48 in the log; UNBIASED (median ratio ~1.00).\n")
cat("  This confirms the v10 document's single 0.448 and resolves it by pair.\n")
cat("* Same region agrees; the divide does not: Copalis<->Mocrocks match at 0.26, but either matches\n")
cat("  Long Beach / Twin Harbors only at ~0.45. Borrowing catch rate across the north-south divide is the worst case.\n")
cat("* Transfer is a SAME-DAY phenomenon: one day off, the error jumps from ~0.45 to ~0.65. Catch rate is set by the\n")
cat("  day (tide, weather) more than the beach, so it is volatile and the day-level piece averages down over a season.\n")
cat("* The HARVEST FLOOR is the season-level component that does NOT average down: ~0.13 (Twin Harbors, matches the\n")
cat("  model's 0.137) to ~0.16-0.18 (Long Beach, higher than the model assumes).\n")
cat(sprintf("* Only measuring the target removes it: borrowing floors the season CPUE error at ~%.2f; interviewing the\n", borrow_floor))
cat(sprintf("  target beach across the season leaves ~%.2f. Donor sampling cannot cross this floor; target interviews can.\n", interview_resid))
cat("  => Spread creel (interview) days across the season at Long Beach / Twin Harbors: that, not more donor days,\n")
cat("     is the lever on harvest uncertainty.\n")
```
