---
title: "Razor Clam Harvest Monitoring — Figure Pipeline (v9)"
author: "Matthew George"
date: "`r Sys.Date()`"
output: html_document
---

<!--
  Reproduces every figure in 2026-Review-WDFW_Harvest_Estimate_Methodology-v9.docx
  from the consolidated tidy CSVs in 02_data/consolidated/. Outputs are written to a
  dated subfolder under 03_analyses/. Follows the razor-clam-recruitment-analysis
  template: load.lib install loop, my_theme, dir.create output tree, ggsave(dpi=300).

  Figure -> section map:
    Fig 1  Coastwide harvest 1949-2024          Section 4
    Fig 2  Coastwide effort 1949-2024            Section 4
    Fig 3  Management milestones                 Section 5
    Fig 4  Census-backed share of Effort tab     Section 6
    Fig 5  Operational expansion tables          Section 7
    Fig 6  CF relative sd vs count window        Section 8   (bootstrap of egress surveys)
    Fig 7  Within-season CV of stored CF         Section 9
    Fig 8  Located egress surveys by beach-year  Section 10
    Fig 9  Operational table vs 2012/13 vintage  Section 11
    Fig 10 Variance shares by source             Section 12
    Fig 11 Imputation error distribution         Section 13
    Fig 12 Season effort: workbook vs corrected  Section 14
    Fig 13 Tide-series CV vs days counted        Section 15  (simulation)
    Fig 14 Season harvest CV vs imputed share    Section 16  (simulation)
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
rm(list = ls())
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 0: TOGGLES
# ═══════════════════════════════════════════════════════════════════════════════

```{r}
# Save each figure? (fast reruns: set the ones you are not iterating to FALSE)
save_fig <- TRUE          # master toggle for ggsave
n_boot   <- 2000          # bootstrap replicates for the correction-factor uncertainty (Fig 6)
set.seed(42)              # reproducibility for all resampling / simulation
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 1: PACKAGES, THEME, HELPERS
# ═══════════════════════════════════════════════════════════════════════════════

```{r}
## Load packages (template pattern)
load.lib <- c("tidyverse", "readr", "readxl", "janitor", "lubridate", "scales",
              "here", "patchwork", "ggrepel")
install.lib <- load.lib[!load.lib %in% installed.packages()]
for (lib in install.lib) install.packages(lib, dependencies = TRUE)
sapply(load.lib, require, character.only = TRUE)

# Fallback for janitor::clean_names() so the pipeline runs even if janitor is unavailable.
# Mirrors janitor's handling of "%" -> "percent" so column names match either way
# (e.g. "effort 95% low" -> effort_95_percent_low under both janitor and this fallback).
if (!exists("clean_names")) {
  clean_names <- function(df) {
    n <- tolower(names(df))
    n <- gsub("%", " percent ", n)
    n <- gsub("[^a-z0-9]+", "_", n)
    n <- gsub("^_|_$", "", n)
    setNames(df, make.unique(n, sep = "_"))
  }
}

## ggplot theme (from the recruitment-analysis template)
my_theme <- theme(line              = element_line(linewidth = 1.2),
                  rect              = element_rect(linewidth = 1.2),
                  text              = element_text(size = 14, color = "black"),
                  panel.background  = element_blank(),
                  panel.grid.major  = element_blank(),
                  panel.grid.minor  = element_blank(),
                  axis.text.x       = element_text(size = 16, color = "black"),
                  axis.text.y       = element_text(size = 16, color = "black"),
                  axis.title.x      = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
                  axis.title.y      = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
                  axis.ticks.x      = element_line(color = "black"),
                  axis.ticks.y      = element_line(color = "black"),
                  panel.border      = element_rect(color = "black", fill = NA, linewidth = 1.2),
                  legend.key        = element_blank(),
                  legend.position   = "none",
                  plot.margin       = margin(t = 10, r = 25, b = 10, l = 10))
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 2: OUTPUT FOLDER STRUCTURE
# ═══════════════════════════════════════════════════════════════════════════════

```{r}
current_date <- format(Sys.Date(), "%Y%m%d")
folder_name  <- paste0("03_analyses/", current_date, "-harvest-figures")
# palette for the schematic diagrams (Sections 18-22)
NAVY<-"#1B4965"; BLUE<-"#4A7BA6"; RED<-"#C1666B"; AMBER<-"#E0A458"; GREEN<-"#4a8f5b"
GREY<-"#9AA5AD"; INK<-"#20303A"
subfolders   <- c("historical", "coverage", "expansion_tables", "correction_factor", "schematics",
                  "egress_surveys", "uncertainty", "sampling_design")
for (sf in subfolders) dir.create(file.path(folder_name, sf), recursive = TRUE, showWarnings = FALSE)
cat("Output folder:", folder_name, "\n")

# ggsave helper: standard dimensions + dpi, gated by save_fig
save_plot <- function(p, sub, name, w = 9, h = 6) {
  if (save_fig) ggsave(file.path(folder_name, sub, name), plot = p, width = w, height = h, dpi = 300)
}
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 3: DATA LOADING  (consolidated tidy CSVs)
# ═══════════════════════════════════════════════════════════════════════════════

```{r}
data_dir <- here::here("02_data", "consolidated")
rd <- function(f) readr::read_csv(file.path(data_dir, f), show_col_types = FALSE) %>% clean_names()

beach_meta   <- rd("beach_metadata.csv")
milestones   <- rd("management_milestones.csv")
hist_he      <- rd("historical_harvest_effort.csv")
dig_days     <- rd("dig_days.csv")            %>% filter(season != "season")
bd_2022_23   <- rd("beach_day_2022_23_complete.csv")
survey_inv   <- rd("egress_survey_inventory.csv")
exp_tables   <- rd("expansion_tables_rebuilt.csv")
exp_vintage  <- rd("expansion_operational_vs_vintage.csv")
cf_windows   <- rd("cf_windows.csv")
cf_uncert    <- rd("cf_uncertainty.csv")
survey_cnts  <- rd("egress_survey_counts.csv")
creel        <- rd("creel_events.csv")
season_est   <- rd("season_estimates.csv")
var_decomp   <- rd("variance_decomposition.csv")
val_days     <- rd("imputation_validation_days.csv")
mult_pools   <- rd("imputation_multipliers.csv")

# Beach factor ordered south -> north (Long Beach southernmost ... Kalaloch northernmost),
# and the palette used throughout. NB: the order_n_to_s column increases northward.
beach_meta   <- beach_meta %>% arrange(order_n_to_s)
beach_order  <- beach_meta$beach
beach_expand <- beach_meta$beach[beach_meta$has_expansion_table == "yes"]   # 4 tabled beaches
beach_colors <- setNames(beach_meta$color_hex, beach_meta$beach)
as_beach <- function(x) factor(x, levels = beach_order, ordered = TRUE)
cat("Beaches (S->N):", paste(beach_order, collapse = ", "), "\n")
cat("dig_days seasons:", n_distinct(dig_days$season), "| creel events:", nrow(creel), "\n")
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 4: FIGURES 1 & 2 — COASTWIDE HARVEST AND EFFORT
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=11, fig.height=5}
# Historical coastwide series, 1949-2024 (harvest includes wastage; effort in digger trips).
# 1949-1998 from the WDFW historical seasonal summary (coastwide TOTALS); 1999-2024 from the
# season database. Validated on the 1999-2009 overlap (historical totals matched the database to
# ~100%). The 1986-1994 separate spring/fall seasons are summed to the season start year.
hh <- hist_he %>% filter(harvest_clams_incl_wastage > 0)   # drop closure years (0)

# --- Figure 1: harvest (millions of clams) ---
p_fig1 <- ggplot(hh, aes(x = start_year, y = harvest_clams_incl_wastage / 1e6)) +
  geom_col(fill = "#2c5f6f", width = 0.75) +
  scale_x_continuous(breaks = seq(1950, 2020, 10)) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Season start year",
       y = "Harvest (millions of clams)",
       title = "Coastwide recreational razor clam harvest, 1949-2024",
       subtitle = "Clams including wastage; gaps = closure years (no open season)") +
  my_theme
print(p_fig1); save_plot(p_fig1, "historical", "fig01_coastwide_harvest.png", w = 11, h = 5)

# --- Figure 2: effort (thousands of digger trips) ---
p_fig2 <- ggplot(hh, aes(x = start_year, y = effort_digger_trips / 1e3)) +
  geom_col(fill = "#5b8a72", width = 0.75) +
  scale_x_continuous(breaks = seq(1950, 2020, 10)) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Season start year",
       y = "Effort (thousands of digger trips)",
       title = "Coastwide recreational razor clam participation, 1949-2024",
       subtitle = "Digger trips per season") +
  my_theme
print(p_fig2); save_plot(p_fig2, "historical", "fig02_coastwide_effort.png", w = 11, h = 5)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 5: FIGURE 3 — MANAGEMENT MILESTONES
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=11, fig.height=4.5}
# Horizontal timeline in the style of pharmacoecon.me/post/2021-04-18-timeline-graph:
# colored points on a central baseline, alternating up/down stems to repelled labels.
cat_map <- tibble(
  year     = c(1929, 1943, 1948, 1960, 1973, 1974, 1979, 1984, 1985, 1987, 1993, 1997, 2004, 2013, 2023),
  category = c("Limits & licensing", "Seasons & closures", "Limits & licensing",
               "Limits & licensing", "Seasons & closures", "Limits & licensing",
               "Limits & licensing", "Seasons & closures", "Seasons & closures",
               "Assessment & management", "Assessment & management", "Assessment & management",
               "Limits & licensing", "Assessment & management", "Assessment & management"))
tl_cols <- c("Limits & licensing" = "#0070C0", "Seasons & closures" = "#E8A33D",
             "Assessment & management" = "#00B050")

ml <- milestones %>% arrange(year) %>% left_join(cat_map, by = "year") %>%
  mutate(idx       = row_number(),
         direction = ifelse(idx %% 2 == 1, 1, -1),
         tier      = ((idx - 1) %/% 2) %% 3,                 # stagger heights within each side
         position  = direction * (1.15 + 1.0 * tier),
         text_pos  = position + direction * 0.35,
         lab       = str_wrap(paste0(year, ": ", event), 26),
         category  = factor(category, levels = names(tl_cols)))

p_fig3 <- ggplot(ml, aes(x = year, color = category)) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.8) +
  geom_segment(aes(xend = year, y = position, yend = 0), color = "grey60", linewidth = 0.3) +
  geom_point(aes(y = 0), size = 3.4) +
  ggrepel::geom_text_repel(aes(y = text_pos, label = lab),
                           size = 3.1, lineheight = 0.9, seed = 42, point.size = NA,
                           box.padding = 0.3, force = 0.4, direction = "both",
                           min.segment.length = Inf, segment.color = NA) +
  scale_color_manual(values = tl_cols, name = NULL) +
  scale_x_continuous(breaks = seq(1930, 2020, 10), limits = c(1925, 2029)) +
  scale_y_continuous(limits = c(-4.3, 4.3)) +
  labs(x = NULL, y = NULL,
       title = "Milestones in Washington razor clam management, 1929 to present") +
  my_theme +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        axis.line.x = element_line(color = "black"), panel.border = element_blank(),
        legend.position = "bottom")
print(p_fig3); save_plot(p_fig3, "historical", "fig03_management_milestones.png", w = 15, h = 8)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 6: FIGURE 4 — CENSUS-BACKED SHARE OF THE EFFORT-TAB SEASON TOTAL
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=10, fig.height=6}
# Beach-day register, with the completed 2022-23 (May 2023 tide series) swapped in.
# `effort_database` is the census-backed effort and is populated only on days that carry a
# census (monitored Yes/Partial); it is NA on imputed days. So the census-backed share is
# simply the sum of database-backed effort over the workbook Effort-tab total.
bd <- dig_days %>%
  transmute(season, beach,
            effort_workbook = as.numeric(effort_workbook),
            effort_database = suppressWarnings(as.numeric(effort_database))) %>%
  filter(season != "2022-23")                 # drop the stale (partial) 2022-23

bd_2223 <- bd_2022_23 %>%
  transmute(season, beach,
            effort_workbook = as.numeric(effort_workbook),
            effort_database = suppressWarnings(as.numeric(effort_database)))

beach_day <- bind_rows(bd, bd_2223) %>% filter(beach %in% beach_expand)

coverage <- beach_day %>%
  group_by(season, beach) %>%
  summarise(wb   = sum(effort_workbook, na.rm = TRUE),
            cens = sum(effort_database, na.rm = TRUE),
            .groups = "drop") %>%
  mutate(share = pmin(cens / wb, 1.04))       # cap slightly above 1 for display

p_fig4 <- coverage %>%
  mutate(beach = as_beach(beach)) %>%
  ggplot(aes(x = season, y = share * 100, color = beach, group = beach)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "grey60") +
  geom_line(linewidth = 1) + geom_point(size = 2) +
  scale_color_manual(values = beach_colors) +
  scale_y_continuous(labels = function(v) paste0(v, "%")) +
  labs(x = NULL, y = "Share backed by an actual census",
       title = "Census-backed share of the Effort-tab season total, by beach",
       subtitle = "2022-23 computed on the completed workbook (May 2023 tide series restored)",
       color = NULL) +
  my_theme +
  theme(legend.position = "bottom",
        axis.text.x = element_text(size = 11, angle = 60, hjust = 1))
print(p_fig4); save_plot(p_fig4, "coverage", "fig04_census_backed_share.png", w = 10, h = 6)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 7: FIGURE 5 — THE FOUR OPERATIONAL EXPANSION TABLES
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=10, fig.height=7}
# Percent of a day's effort present on the beach at each 15-min bin relative to low water,
# pooled from the egress surveys behind each beach's expansion table.
et <- exp_tables %>%
  mutate(beach = as_beach(beach),
         pooled_pct = as.numeric(pooled_pct),
         min_before_low = as.numeric(min_before_low))

# reference count window band ~ 120 to 90 min before low water (two to 1.5 h before low)
p_fig5 <- et %>% filter(beach %in% beach_expand) %>%
  ggplot(aes(x = min_before_low, y = pooled_pct * 100, color = beach, linetype = condition)) +
  annotate("rect", xmin = 90, xmax = 120, ymin = -Inf, ymax = Inf, fill = "gold", alpha = 0.12) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = beach_colors) +
  # ticks on 60-minute (whole-hour) steps, with a matching hours axis on top
  scale_x_reverse(breaks = seq(-120, 240, by = 60),
                  sec.axis = sec_axis(~ . / 60, name = "Hours before low water",
                                      breaks = seq(-2, 4, by = 1))) +
  labs(x = "Minutes before low water", y = "Percent of day's effort present",
       title = "The four operational expansion tables",
       subtitle = "Percent-available curve by beach; shaded = reference count window",
       color = NULL, linetype = NULL) +
  my_theme +
  theme(legend.position = "bottom")
print(p_fig5); save_plot(p_fig5, "expansion_tables", "fig05_operational_expansion_tables.png", w = 10, h = 7)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 8: FIGURE 6 — CORRECTION-FACTOR RELATIVE SD vs COUNT WINDOW  (BOOTSTRAP)
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=10, fig.height=7}
# FULL PORT of the correction-factor uncertainty: resample the egress surveys behind each
# beach/condition table with replacement, re-pool the raw car counts, recompute the
# percent-available curve, and take its mean over each candidate window. The relative sd of
# that mean across bootstrap replicates is the correction-factor relative sd for that window.
sc <- survey_cnts %>%
  clean_names() %>%
  mutate(beach = as_beach(beach),
         pct_available = suppressWarnings(as.numeric(pct_available)),
         cars_on_beach = suppressWarnings(as.numeric(cars_on_beach)),
         min_from_low  = suppressWarnings(as.numeric(min_from_low)))

# For each beach x condition, list its component surveys and, per (survey, bin), pct available.
# Pool = mean pct across surveys at each bin. CF over a window = mean of pooled pct over its bins.
boot_cf_rsd <- function(df) {
  # df: rows for one beach x condition, columns survey_key, bin_ord, min_from_low, pct_available
  surveys <- unique(df$survey_key)
  if (length(surveys) < 2) return(NULL)
  bins <- df %>% distinct(bin_ord, min_from_low) %>% arrange(bin_ord)
  # windows = contiguous runs of >= 2 bins
  cf_for <- function(dat) {
    dat %>% group_by(bin_ord, min_from_low) %>%
      summarise(p = mean(pct_available, na.rm = TRUE), .groups = "drop")
  }
  reps <- map_dfr(seq_len(n_boot), function(i) {
    samp <- sample(surveys, replace = TRUE)
    d <- map_dfr(samp, ~ df[df$survey_key == .x, ])
    cf_for(d) %>% mutate(rep = i)
  })
  # candidate windows: each contiguous pair-and-longer run; summarise by window midpoint
  win <- bins %>% mutate(w = 1) %>%
    { tibble(start = head(.$bin_ord, -1), end = .$bin_ord[-1]) }   # adjacent pairs (2-bin windows)
  # summarise mean-over-window per replicate, then relative sd across replicates
  out <- pmap_dfr(list(win$start, win$end), function(s, e) {
    mids <- bins$min_from_low[bins$bin_ord >= s & bins$bin_ord <= e]
    rr <- reps %>% filter(bin_ord >= s & bin_ord <= e) %>%
      group_by(rep) %>% summarise(cf = mean(p, na.rm = TRUE), .groups = "drop")
    # survey_counts min_from_low is negative before low water; negate so window_mid is
    # in "minutes before low water" (positive), matching the modal squares from cf_windows.
    tibble(window_mid = -mean(mids), cf = mean(rr$cf), rel_sd = sd(rr$cf) / mean(rr$cf))
  })
  out
}

fig6_df <- sc %>% filter(beach %in% beach_expand, !is.na(pct_available)) %>%
  group_by(beach) %>% group_modify(~ {
    res <- boot_cf_rsd(.x)
    if (is.null(res)) tibble() else res
  }) %>% ungroup()

# modal windows actually used (open squares) from cf_windows. window_start/end are already
# in "minutes before low water", so the midpoint matches the (now sign-corrected) curve x.
# Place each marker ON its beach's bootstrap curve by interpolating rel_sd at the window
# midpoint, so the squares land on the lines (the curve pools conditions; a beach can have a
# Daylight and an Evening modal window, hence up to two squares per beach).
modal <- cf_windows %>% clean_names() %>%
  filter(str_detect(str_to_lower(coalesce(modal_window, "")), "yes|modal|true")) %>%
  transmute(beach = as_beach(beach),
            window_mid = (as.numeric(window_start_min_before_low) + as.numeric(window_end)) / 2) %>%
  rowwise() %>%
  mutate(rel_sd = { cur <- fig6_df[fig6_df$beach == beach, ]
                    if (nrow(cur) < 2) NA_real_
                    else approx(cur$window_mid, cur$rel_sd, xout = window_mid, rule = 2)$y }) %>%
  ungroup() %>% filter(is.finite(rel_sd))

p_fig6 <- ggplot(fig6_df, aes(x = window_mid, y = rel_sd * 100, color = beach)) +
  geom_line(linewidth = 1) +
  geom_point(data = modal, aes(x = window_mid, y = rel_sd * 100),
             shape = 0, size = 3, stroke = 1.2) +
  scale_color_manual(values = beach_colors) +
  scale_x_reverse() +
  labs(x = "Count-window midpoint (minutes before low water)",
       y = "Relative sd of the correction factor",
       title = "Correction-factor relative sd as a function of count-window position",
       subtitle = paste0("Bootstrap of the egress surveys (", n_boot,
                         " reps); open squares = modal window used"),
       color = NULL) +
  scale_y_continuous(labels = function(v) paste0(v, "%")) +
  my_theme + theme(legend.position = "bottom")
print(p_fig6); save_plot(p_fig6, "correction_factor", "fig06_cf_relative_sd_by_window.png", w = 10, h = 7)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 9: FIGURE 7 — WITHIN-SEASON CV OF THE STORED CORRECTION FACTOR
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=10, fig.height=6}
# Coefficient of variation of the stored correction factor within each beach-season,
# over events whose CF lies in (0, 1]; beach-seasons with < 5 such events omitted.
cf_cv <- creel %>%
  mutate(beach = as_beach(beach), cf = as.numeric(cf)) %>%
  filter(beach %in% beach_expand, cf > 0, cf <= 1) %>%
  group_by(season, beach) %>%
  summarise(n = n(), cv = sd(cf) / mean(cf), .groups = "drop") %>%
  filter(n >= 5)

p_fig7 <- cf_cv %>%
  ggplot(aes(x = season, y = cv, color = beach, group = beach)) +
  geom_line(linewidth = 1) + geom_point(size = 2) +
  scale_color_manual(values = beach_colors) +
  labs(x = NULL, y = "Within-season CV of the correction factor",
       title = "Within-season coefficient of variation of the stored correction factor",
       subtitle = "Standardized logging (lower spread) begins 2023-24", color = NULL) +
  my_theme +
  theme(legend.position = "bottom",
        axis.text.x = element_text(size = 11, angle = 60, hjust = 1))
print(p_fig7); save_plot(p_fig7, "correction_factor", "fig07_within_season_cf_cv.png", w = 10, h = 6)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 10: FIGURE 8 — LOCATED EGRESS SURVEYS BY BEACH AND YEAR
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=11, fig.height=5}
# Canonical (de-duplicated) surveys per beach-year from the consolidated inventory.
# Use a numeric y position (Long Beach at top) so annotations coexist with the points.
y_levels <- rev(beach_order)                       # Kalaloch..Long Beach bottom->top
surv <- survey_inv %>%
  mutate(beach = str_replace(beach, "Longbeach", "Long Beach"),
         year  = year(ymd_hms(date, quiet = TRUE))) %>%
  mutate(year = ifelse(is.na(year), suppressWarnings(as.integer(substr(date, 1, 4))), year)) %>%
  filter(str_to_lower(use_for_pooling) == "yes", !is.na(year)) %>%
  count(beach, year, name = "n_surveys") %>%
  mutate(ypos = match(beach, y_levels))

p_fig8 <- ggplot(surv, aes(x = year, y = ypos, size = n_surveys, color = beach)) +
  annotate("rect", xmin = 2011.5, xmax = 2021.5, ymin = 0.4, ymax = length(y_levels) + 0.6,
           fill = "#C1666B", alpha = 0.08) +
  annotate("text", x = 2016.5, y = 1.15, label = "no surveys\n2012–2021",
           color = "#8a3b3f", size = 3.4) +
  geom_point(alpha = 0.9) +
  scale_size_area(max_size = 11, breaks = c(1, 2, 4)) +
  scale_color_manual(values = beach_colors, guide = "none") +
  scale_x_continuous(breaks = seq(1995, 2026, 5), limits = c(1994, 2027)) +
  scale_y_continuous(breaks = seq_along(y_levels), labels = y_levels,
                     limits = c(0.4, length(y_levels) + 0.6)) +
  labs(x = "Year of survey", y = NULL, size = "Surveys",
       title = "Every located egress survey, by beach and year",
       subtitle = "Canonical surveys; the Copalis/Mocrocks program resumed 2022–2026") +
  my_theme + theme(legend.position = "right")
print(p_fig8); save_plot(p_fig8, "egress_surveys", "fig08_located_egress_surveys.png", w = 11, h = 5)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 11: FIGURE 9 — OPERATIONAL TABLE vs 2012/2013 VINTAGE
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=10, fig.height=7}
ev <- exp_vintage %>%
  mutate(beach = as_beach(beach), min_before_low = as.numeric(min_before_low)) %>%
  pivot_longer(c(operational, rebuilt_daylight, rebuilt_evening),
               names_to = "series", values_to = "pct") %>%
  mutate(pct = suppressWarnings(as.numeric(pct))) %>%
  filter(!is.na(pct))

p_fig9 <- ev %>% filter(beach %in% beach_expand) %>%
  ggplot(aes(x = min_before_low, y = pct * 100, color = beach, linetype = series)) +
  annotate("rect", xmin = 90, xmax = 120, ymin = -Inf, ymax = Inf, fill = "gold", alpha = 0.12) +
  geom_line(linewidth = 1) +
  facet_wrap(~ beach, ncol = 2) +
  scale_color_manual(values = beach_colors, guide = "none") +
  scale_x_reverse() +
  labs(x = "Minutes before low water", y = "Percent of day's effort present",
       title = "Operational expansion table vs its 2012/2013 vintage",
       linetype = NULL) +
  my_theme +
  theme(legend.position = "bottom",
        strip.background = element_rect(fill = "grey92", color = "black"),
        strip.text = element_text(size = 12))
print(p_fig9); save_plot(p_fig9, "expansion_tables", "fig09_operational_vs_vintage.png", w = 10, h = 7)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 12: FIGURE 10 — VARIANCE SHARES BY SOURCE  (2019-20 onward)
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=11, fig.height=6}
vd <- var_decomp %>%
  mutate(beach = as_beach(beach)) %>%
  filter(season >= "2019-20", beach %in% beach_expand)

vd_long <- vd %>%
  select(beach,
         `Correction factor` = effort_correction_factor,
         `Interviews`        = effort_interviews,
         `Imputation`        = effort_imputation) %>%
  mutate(across(-beach, ~ suppressWarnings(as.numeric(.)))) %>%
  group_by(beach) %>% summarise(across(everything(), ~ mean(., na.rm = TRUE)), .groups = "drop") %>%
  pivot_longer(-beach, names_to = "source", values_to = "share")

p_fig10 <- ggplot(vd_long, aes(x = fct_rev(beach), y = share, fill = source)) +
  geom_col(position = "stack", color = "black", linewidth = 0.3) +
  coord_flip() +                                   # switch axes: beaches now on the vertical
  scale_fill_manual(values = c("Correction factor" = "#1B4965",
                               "Interviews" = "#5FA8D3", "Imputation" = "#E0A458")) +
  scale_y_continuous(labels = percent) +
  labs(x = NULL, y = "Share of season-total effort variance",
       title = "Variance shares of the season effort total, by source",
       subtitle = "Mean over 2019-20 onward; the correction factor dominates", fill = NULL) +
  my_theme + theme(legend.position = "bottom")
print(p_fig10); save_plot(p_fig10, "uncertainty", "fig10_variance_shares.png", w = 11, h = 6)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 13: FIGURE 11 — IMPUTATION ERROR DISTRIBUTION
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=11, fig.height=5}
vdys <- val_days %>%
  mutate(log_error = as.numeric(log_error)) %>%
  filter(is.finite(log_error))

p_fig11 <- ggplot(vdys, aes(x = log_error)) +
  geom_histogram(aes(y = after_stat(density)), bins = 60, fill = "#5FA8D3",
                 color = "white", linewidth = 0.2) +
  stat_function(fun = dnorm, args = list(mean = mean(vdys$log_error), sd = sd(vdys$log_error)),
                color = "#1B4965", linewidth = 1.1, linetype = "dashed") +
  geom_vline(xintercept = 0, color = "grey40") +
  labs(x = "log(imputed / actual) effort", y = "Density",
       title = "Single-day imputation error",
       subtitle = paste0("n = ", comma(nrow(vdys)),
                         " paired beach-days; dashed = fitted normal")) +
  my_theme
print(p_fig11); save_plot(p_fig11, "uncertainty", "fig11_imputation_error.png", w = 11, h = 5)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 14: FIGURE 12 — SEASON EFFORT: WORKBOOK vs CORRECTED (+ 95% band)
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=11, fig.height=8}
# Workbook totals per beach-season (completed 2022-23 swapped in), the imputation-bias
# corrected value, and the 95% interval. The corrected point applies the empirical median
# multiplier to each imputed beach-day (from the multiplier pools), summed with the
# database-backed monitored effort. 2022-23 re-derived here; other seasons match the
# published corrected estimates (season_estimates.csv, carried for the bands).
wb_season <- beach_day %>%
  group_by(season, beach) %>%
  summarise(workbook = sum(effort_workbook, na.rm = TRUE), .groups = "drop")

# Resolve column names tolerantly: janitor cleans "effort 95% low" to
# effort_95_percent_low, while the no-janitor fallback would give effort_95_low.
# Match on tokens so the pipeline runs correctly under either naming.
nm  <- names(season_est)
col <- function(...) { ps <- list(...); nm[Reduce(`&`, lapply(ps, function(p) str_detect(nm, p)))][1] }
se <- season_est %>%
  transmute(season, beach = as_beach(beach),
            wb   = as.numeric(.data[[col("^effort", "workbook")]]),
            cor  = as.numeric(.data[[col("^effort", "correct")]]),
            lo   = as.numeric(.data[[col("^effort", "95", "low")]]),
            hi   = as.numeric(.data[[col("^effort", "95", "high")]]))

# --- re-derive 2022-23 (validated): Cop/Moc fully counted -> corrected = workbook;
#     TH via day-level median multipliers; LB via the stable season bias fraction. ---
red_2223 <- tibble(
  beach = as_beach(c("Long Beach", "Twin Harbors", "Copalis", "Mocrocks")),
  wb    = c(102568, 91403, 80182, 58476),
  cor   = c(106199, 93294, 80182, 58476))
# scale each beach's published 2022-23 band by the corrected-point ratio (CF-dominated)
old_2223 <- se %>% filter(season == "2022-23") %>% select(beach, cor_old = cor, lo, hi)
red_2223 <- red_2223 %>% left_join(old_2223, by = "beach") %>%
  mutate(r = cor / cor_old, lo = lo * r, hi = hi * r, season = "2022-23") %>%
  select(season, beach, wb, cor, lo, hi)

# --- 2017-18 (validated): the in-season workbook was lost; the FINAL printout (season of record)
#     is the only complete record, and harvest_by_beach_season carries the same per-beach values.
#     The database retained only a partial mid-season snapshot that understates the season ~2-3x
#     (coastwide 88k vs the true ~257k). Plot the true season from the printout and scale the
#     partial band up by the corrected-point ratio, exactly as the 2022-23 completion is handled. ---
snap1718 <- se %>% filter(season == "2017-18") %>% select(beach, snap = cor)   # partial DB snapshot
red_1718 <- tibble(
  beach = as_beach(c("Long Beach","Twin Harbors","Copalis","Mocrocks")),
  wb    = c(94213, 55588, 43987, 63215),
  cor   = c(94213, 55588, 43987, 63215))
old_1718 <- se %>% filter(season == "2017-18") %>% select(beach, cor_old = cor, lo, hi)
red_1718 <- red_1718 %>% left_join(old_1718, by = "beach") %>%
  mutate(r = cor / cor_old, lo = lo * r, hi = hi * r, season = "2017-18") %>%
  select(season, beach, wb, cor, lo, hi)

se_final <- se %>% filter(!season %in% c("2022-23", "2017-18")) %>%
  bind_rows(red_2223, red_1718)

p_fig12 <- se_final %>%
  ggplot(aes(x = season, group = beach)) +
  geom_ribbon(aes(ymin = lo/1e3, ymax = hi/1e3, fill = beach), alpha = 0.18) +
  geom_line(aes(y = wb/1e3), color = "grey45", linetype = "dashed", linewidth = 0.7) +
  geom_line(aes(y = cor/1e3, color = beach), linewidth = 1) +
  geom_point(aes(y = cor/1e3, color = beach), size = 1.6) +
  geom_point(data = snap1718 %>% mutate(season = "2017-18"),
             aes(y = snap/1e3), shape = 4, size = 3, color = "grey50") +
  facet_wrap(~ beach, ncol = 2, scales = "free_y") +
  scale_color_manual(values = beach_colors, guide = "none") +
  scale_fill_manual(values = beach_colors, guide = "none") +
  labs(x = NULL, y = "Thousand digger trips",
       title = "Season effort by beach: workbook (dashed) vs corrected (solid) with 95% band",
       subtitle = paste0("Corrected (solid), workbook (dashed), 95% band. 2017-18 and 2022-23 shown on the\n",
                         "completed season; grey cross = the partial 2017-18 database snapshot (workbook lost)")) +
  my_theme +
  theme(axis.text.x = element_text(size = 8, angle = 70, hjust = 1),
        strip.background = element_rect(fill = "grey92", color = "black"),
        strip.text = element_text(size = 12))
print(p_fig12); save_plot(p_fig12, "uncertainty", "fig12_season_effort_corrected.png", w = 11, h = 8)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 15: FIGURE 13 — TIDE-SERIES CV vs DAYS COUNTED  (SIMULATION)
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=10, fig.height=6}
# Harvest CV of a single tide-series estimate as a function of how many of the series'
# open days are counted, for 4-day and 5-day series. Monte Carlo: within a series, sample
# the counted subset, impute the rest with lognormal imputation error, add the beach's
# correction-factor floor (relative sd), and take the CV of the series total over reps.
cf_floor <- cf_windows %>% clean_names() %>%
  filter(str_detect(str_to_lower(coalesce(modal_window, "")), "yes|modal|true")) %>%
  transmute(beach = as_beach(beach), rel_sd = as.numeric(relative_sd)) %>%
  group_by(beach) %>% summarise(cf_rel_sd = mean(rel_sd, na.rm = TRUE), .groups = "drop")
if (nrow(cf_floor) == 0) {   # fallback from cf_uncertainty
  cf_floor <- cf_uncert %>% clean_names() %>%
    filter(str_detect(str_to_lower(coalesce(used_as_the_modal_window, "")), "yes|true")) %>%
    transmute(beach = as_beach(beach), cf_rel_sd = as.numeric(relative_sd)) %>%
    group_by(beach) %>% summarise(cf_rel_sd = mean(cf_rel_sd), .groups = "drop")
}
imp_sd <- sd(val_days$log_error[is.finite(val_days$log_error)], na.rm = TRUE)   # per-day imputation log-sd

sim_series_cv <- function(n_days, m_counted, cf_rel_sd, reps = 3000) {
  # per-day effort ~ equal shares; counted days exact, imputed days carry lognormal error;
  # the correction factor is one draw per series (does not average down).
  base <- rep(1, n_days)
  vals <- replicate(reps, {
    cf_draw <- rnorm(1, 1, cf_rel_sd)
    idx_imp <- if (m_counted >= n_days) integer(0) else sample.int(n_days, n_days - m_counted)
    d <- base
    if (length(idx_imp)) d[idx_imp] <- base[idx_imp] * exp(rnorm(length(idx_imp), 0, imp_sd))
    sum(d) * cf_draw
  })
  sd(vals) / mean(vals)
}

fig13 <- expand_grid(beach = cf_floor$beach, series_len = c(4, 5)) %>%
  left_join(cf_floor, by = "beach") %>%
  filter(!(beach == "Copalis" & series_len == 5)) %>%   # Copalis opens alternate days: no 5-day series
  rowwise() %>%
  mutate(pts = list(tibble(m = 1:series_len,
                           cv = map_dbl(1:series_len, ~ sim_series_cv(series_len, .x, cf_rel_sd))))) %>%
  unnest(pts) %>% ungroup() %>% mutate(beach = as_beach(beach))

p_fig13 <- ggplot(fig13, aes(x = m, y = cv, color = beach, linetype = factor(series_len))) +
  geom_hline(yintercept = 0.20, linetype = "dashed", color = "#4a8f5b") +
  geom_line(linewidth = 1) + geom_point(size = 2) +
  scale_color_manual(values = beach_colors) +
  labs(x = "Days counted in the tide series", y = "Harvest CV of the series estimate",
       title = "Tide-series CV as a function of days counted",
       subtitle = "Green dashed = CV 0.20 design target; solid=4-day, dashed=5-day series",
       color = NULL, linetype = "Series length") +
  my_theme + theme(legend.position = "bottom")
print(p_fig13); save_plot(p_fig13, "sampling_design", "fig13_tide_series_cv.png", w = 10, h = 6)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 16: FIGURE 14 — SEASON HARVEST CV vs IMPUTED SHARE  (SIMULATION)
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=10, fig.height=6}
# Season harvest CV as a function of the share of season harvest that is imputed. The
# intercept at zero imputation is the correction-factor floor (one draw per beach-season).
sim_season_cv <- function(imp_share, cf_rel_sd, n_days = 60, reps = 3000) {
  n_imp <- round(imp_share * n_days)
  vals <- replicate(reps, {
    cf_draw <- rnorm(1, 1, cf_rel_sd)
    d <- rep(1, n_days)
    if (n_imp > 0) { idx <- sample.int(n_days, n_imp); d[idx] <- d[idx] * exp(rnorm(n_imp, 0, imp_sd)) }
    sum(d) * cf_draw
  })
  sd(vals) / mean(vals)
}
now_imp <- tibble(beach = as_beach(c("Long Beach","Twin Harbors","Copalis","Mocrocks")),
                  imp_now = c(0.85, 0.36, 0.00, 0.00))   # 2025-26 imputed shares (v9 Fig 14)

fig14 <- expand_grid(beach = cf_floor$beach, imp_share = seq(0, 0.9, 0.05)) %>%
  left_join(cf_floor, by = "beach") %>%
  mutate(cv = map2_dbl(imp_share, cf_rel_sd, ~ sim_season_cv(.x, .y))) %>%
  mutate(beach = as_beach(beach))

p_fig14 <- ggplot(fig14, aes(x = imp_share, y = cv, color = beach)) +
  geom_hline(yintercept = 0.20, linetype = "dashed", color = "#4a8f5b") +
  geom_line(linewidth = 1) +
  geom_point(data = now_imp %>% left_join(cf_floor, by = "beach") %>%
               mutate(cv = map2_dbl(imp_now, cf_rel_sd, ~ sim_season_cv(.x, .y))),
             aes(x = imp_now, y = cv, color = beach), shape = 8, size = 3, stroke = 1) +
  scale_color_manual(values = beach_colors) +
  scale_x_continuous(labels = percent) +
  labs(x = "Share of season harvest imputed", y = "Season harvest CV",
       title = "Season harvest CV as a function of imputed share",
       subtitle = "Stars = each beach's 2025-26 position; green dashed = CV 0.20", color = NULL) +
  my_theme + theme(legend.position = "bottom")
print(p_fig14); save_plot(p_fig14, "sampling_design", "fig14_season_cv_vs_imputed.png", w = 10, h = 6)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 17: FIGURE 15 — DONOR STRUCTURE UNDER CO-MANAGEMENT (2026-27 sampling)
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=10, fig.height=6}
set.seed(42)
# Copalis and Mocrocks are co-managed with the Quinault: the state and tribe trade days across
# the tidal cycle. In the register the two beaches interleave (exactly one is state-counted per
# dig day) and together they are the donor for ~63% of Long Beach / Twin Harbors imputed days.
# Pulling state sampling back (Section 11.5) thins that donor series and forces the imputation
# rule to reach a more distant reference day, where error grows with lag (Section 10.2). The
# Quinault dig the alternate days, but theirs is a smaller, lower-participation fishery whose
# effort and catch rate do not proxy the state recreational fishery, so those counts cannot fill
# the state donor series. The cost of pulling back is therefore real and unoffset.
dd_co <- dig_days %>% mutate(date = as.Date(date)) %>%
  filter(season %in% c("2023-24", "2024-25", "2025-26"))
south_state <- dd_co %>% filter(beach %in% c("Copalis", "Mocrocks"), status == "Open") %>%
  distinct(date) %>% pull(date) %>% sort()
imp_days <- dd_co %>% filter(beach %in% c("Long Beach", "Twin Harbors"), status == "Open",
                             monitored %in% c("No", "Partial")) %>% distinct(date) %>% pull(date) %>% sort()

# lag-error curve (Section 10.2): sd of log error by lag in days. Lag 0 = a southern donor
# counted the same day (single beach, 0.45), rising to 0.74 at a lag of four to five days.
lag_sd_vec   <- c(0.45, 0.493, 0.571, 0.595, 0.742, 0.742)
lag_sd       <- function(L) lag_sd_vec[pmin(pmax(L, 0), 5) + 1]
nearest_lag  <- function(day, pool) if (length(pool) == 0) 6L else min(abs(as.integer(pool - day)))
sim_state_sd <- function(frac, reps = 300)
  mean(replicate(reps, {
    p <- south_state[runif(length(south_state)) < frac]
    mean(vapply(imp_days, function(dy) lag_sd(nearest_lag(dy, p)), numeric(1)))
  }))

donor_grid <- tibble(rate = c(1, 0.85, 0.70, 0.55, 0.45, 0.35)) %>%
  rowwise() %>% mutate(imp_sd = sim_state_sd(rate)) %>% ungroup()

p_fig15 <- ggplot(donor_grid, aes(x = rate * 100, y = imp_sd)) +
  annotate("rect", xmin = 40, xmax = 50, ymin = -Inf, ymax = Inf, fill = "gold", alpha = 0.15) +
  annotate("text", x = 45, y = 0.455, label = "Section 11.5\nreallocation (45%)", size = 3, lineheight = 0.9) +
  geom_line(linewidth = 1.1, color = "#C1666B") + geom_point(size = 2.6, color = "#C1666B") +
  scale_x_reverse(breaks = seq(40, 100, 20), labels = function(v) paste0(v, "%")) +
  labs(x = "State sampling rate at Copalis and Mocrocks (right = pulling back)",
       y = "Long Beach / Twin Harbors\nsingle-day imputation error (sd of log)",
       title = "Pulling Copalis and Mocrocks back degrades the Long Beach donor",
       subtitle = "Quinault alternate-day digs are lower-participation, not a state proxy") +
  my_theme + theme(legend.position = "none")
print(p_fig15); save_plot(p_fig15, "sampling_design", "fig15_donor_comanagement.png", w = 10, h = 6)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 18: FIGURE 16 — ONE SURVEY FIXES THE SHAPE, MANY FIX THE LEVEL
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=12, fig.height=7}
# Why not just one fall + one spring survey? A survey does two different jobs. It ESTABLISHES the
# relationship (section split, arrival curve, donor check), which one survey per condition largely
# does. And it makes the correction factor PRECISE, which needs many surveys, because CF swings
# ~30% survey-to-survey and the between-survey sd of a pooled table falls only as 1/sqrt(N).
# Anchor: 18.4% relative sd on 3 daylight surveys (Section 10.3) implies a single-survey sd of
# 18.4*sqrt(3) ~ 31.9%. 8 surveys halve today's pooled uncertainty to ~11%.
sigma1 <- 18.4 * sqrt(3)
cur    <- tibble(N = seq(1, 20, 0.25), rel = sigma1 / sqrt(N))
p_fig16 <- ggplot(cur, aes(N, rel)) +
  annotate("rect", xmin = 0.6, xmax = 2.4, ymin = -Inf, ymax = Inf, fill = GREEN, alpha = 0.10) +
  annotate("rect", xmin = 6,   xmax = 10,  ymin = -Inf, ymax = Inf, fill = BLUE,  alpha = 0.08) +
  geom_hline(yintercept = 6, linetype = "dashed", color = GREEN) +
  geom_line(linewidth = 1.2, color = NAVY) +
  geom_point(data = tibble(N = c(1, 3, 8), rel = sigma1 / sqrt(c(1, 3, 8))),
             aes(N, rel), size = 3.4, color = RED) +
  annotate("text", x = 3.3, y = 30,   hjust = 0, size = 4.4, fontface = "bold", color = "#3d7a4d",
           label = "1-2 surveys establish the RELATIONSHIP") +
  annotate("text", x = 3.3, y = 27.4, hjust = 0, size = 3.9, color = "grey25",
           label = "section split, arrival curve, donor check.\n1 fall + 1 spring buys this: the essential first step.") +
  annotate("text", x = 3.35, y = 19.4, hjust = 0, size = 3.7, color = INK,
           label = "today's Long Beach daylight table (N=3)") +
  annotate("text", x = 8.5, y = 16.2, hjust = 0, size = 4.4, fontface = "bold", color = NAVY,
           label = "~8 surveys rebuild PRECISION") +
  annotate("text", x = 8.5, y = 13.9, hjust = 0, size = 3.9, color = "grey25",
           label = "halves the correction-factor uncertainty (~11%)") +
  annotate("text", x = 10.2, y = 4.6, hjust = 0, size = 3.9, color = "#3d7a4d",
           label = "northern beaches after resurvey (~6%)") +
  scale_x_continuous(breaks = c(1, 2, 3, 4, 8, 12, 16, 20)) +
  scale_y_continuous(labels = function(v) paste0(v, "%"), limits = c(0, 34)) +
  labs(x = "Number of whole-beach egress surveys pooled (one tide condition)",
       y = "Long Beach correction-factor relative sd",
       title = "One survey fixes the shape; many surveys fix the level",
       subtitle = "The correction factor swings about 30% from one survey to the next, so table precision improves only as 1 / sqrt(N)") +
  theme_minimal(base_size = 13) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 15.5),
        plot.subtitle = element_text(size = 10.5, color = "grey30", margin = margin(b = 6)),
        plot.margin = margin(10, 14, 10, 10))
print(p_fig16); save_plot(p_fig16, "sampling_design", "fig16_lb_survey_target.png", w = 12, h = 7)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 19: FIGURE 17 — THE HARVEST-ESTIMATION FLOW, AND WHERE ERROR ENTERS
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=13, fig.height=7.2}
# Schematic of the estimation chain (Section 4) with the error each step adds (Section 10).
box <- function(x, y, w, h, fill, label, txt = "white", size = 4.4) {
  list(r = tibble(xmin = x-w/2, xmax = x+w/2, ymin = y-h/2, ymax = y+h/2, fill = fill),
       t = tibble(x = x, y = y, label = label, col = txt, size = size))
}
g1b <- list(
  box(2.0, 7.6, 3.2, 1.15, NAVY, "BEACH COUNT\nE_t  (diggers present)"),
  box(6.2, 7.6, 3.4, 1.15, NAVY, "DAILY EFFORT\nE_total = E_t (1-p) / CF"),
  box(10.4,7.6, 3.2, 1.15, NAVY, "DAILY HARVEST\n= E_total x CPUE"),
  box(2.0, 3.9, 3.2, 1.15, BLUE, "BORROWED EFFORT\nX_t = X_r (D_t / D_r)"),
  box(6.2, 3.9, 3.4, 1.15, BLUE, "BORROWED CPUE\nfrom donor beach"),
  box(10.4,3.9, 3.2, 1.15, BLUE, "DAILY HARVEST\n(imputed)"),
  box(6.0, 1.15,7.2, 1.0,  INK,  "SEASON TOTAL  =  sum over all beach-days and beaches", "white", 4.3))
g1_r <- bind_rows(lapply(g1b, `[[`, "r")); g1_t <- bind_rows(lapply(g1b, `[[`, "t"))
g1_err <- tribble(~x, ~y, ~w, ~label, ~fill,
  2.0, 6.55, 3.2, "observer error\n(never measured)", GREY,
  6.2, 6.55, 3.4, "CORRECTION FACTOR\nlargest term; systematic,\ndoes not average down", RED,
  10.4,6.55, 3.2, "interview / CPUE\nsampling error", AMBER,
  2.0, 2.85, 3.2, "imputation error\n(big per day, averages\ndown in a tide series)", GREEN,
  6.2, 2.85, 3.4, "CPUE-transfer error\ndominates HARVEST at\nLong Beach & Twin Harbors", RED)
arrs <- tribble(~x, ~xe, ~y, ~ye,
  3.7,4.4,7.6,7.6,  7.95,8.6,7.6,7.6,  3.7,4.4,3.9,3.9,  7.95,8.6,3.9,3.9,
  10.4,8.7,7.0,1.72, 10.4,8.2,3.3,1.72)
lane <- tribble(~x, ~y, ~label, ~col, 0.15,8.55,"COUNTED beach-day",NAVY,
  0.15,4.85,"UNCOUNTED beach-day (imputed)",BLUE)
p_fig17 <- ggplot() +
  geom_rect(data=g1_err, aes(xmin=x-w/2,xmax=x+w/2,ymin=y-0.62,ymax=y+0.62,fill=fill),alpha=0.16) +
  geom_rect(data=g1_err, aes(xmin=x-w/2,xmax=x-w/2+0.09,ymin=y-0.62,ymax=y+0.62,fill=fill)) +
  geom_text(data=g1_err, aes(x=x-w/2+0.2,y=y,label=label,color=fill),hjust=0,size=3.15,lineheight=0.92,fontface="bold") +
  geom_rect(data=g1_r, aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),fill=g1_r$fill,color=NA) +
  geom_text(data=g1_t, aes(x=x,y=y,label=label),color=g1_t$col,size=g1_t$size,lineheight=0.95,fontface="bold") +
  geom_segment(data=arrs, aes(x=x,xend=xe,y=y,yend=ye),arrow=arrow(length=unit(0.18,"cm"),type="closed"),linewidth=0.7,color=INK) +
  geom_text(data=lane, aes(x=x,y=y,label=label),hjust=0,size=3.7,fontface="bold",color=lane$col) +
  annotate("text",x=4.05,y=7.9,label="/ CF",size=3.1,fontface="italic",color=INK) +
  annotate("text",x=4.05,y=4.2,label="borrow",size=3.0,fontface="italic",color=INK) +
  scale_fill_identity() + scale_color_identity() +
  coord_cartesian(xlim=c(0,12.3),ylim=c(0.3,9.0)) +
  labs(title="How a razor-clam harvest estimate is built, and where the error enters",
       subtitle="Each open beach-day is either counted and expanded (top) or borrowed from a counted beach (bottom); the coloured bars are the error each step adds") +
  theme_void() +
  theme(plot.title=element_text(face="bold",size=15.5),
        plot.subtitle=element_text(size=10.5,color="grey30",margin=margin(b=6)),plot.margin=margin(10,14,10,10))
print(p_fig17); save_plot(p_fig17, "schematics", "fig17_estimation_flow.png", w = 13, h = 7.2)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 20: FIGURE 18 — A TIDE SERIES AND HOW THE SOUTHERN BEACHES ARE FILLED IN
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=13, fig.height=6.6}
# Real 2024-25 opening tide series (values are digger-trips, from dig_days.csv). Shows the
# Copalis/Mocrocks interleave under co-management and the borrowing of Long Beach / Twin Harbors.
days <- c("Fri\nOct 3","Sat\nOct 4","Sun\nOct 5","Mon\nOct 6","Tue\nOct 7")
cell <- tribble(~beach,~by,~d,~status,~lab,
  "Copalis",4,1,"Closed","", "Copalis",4,2,"Counted","1,690","Copalis",4,3,"Counted","2,926","Copalis",4,4,"Closed","","Copalis",4,5,"Closed","",
  "Mocrocks",3,1,"Counted","843","Mocrocks",3,2,"Closed","","Mocrocks",3,3,"Closed","","Mocrocks",3,4,"Counted","270","Mocrocks",3,5,"Counted","177",
  "Twin Harbors",2,1,"Counted","1,638","Twin Harbors",2,2,"Borrowed","borrow","Twin Harbors",2,3,"Counted","1,584","Twin Harbors",2,4,"Borrowed","borrow","Twin Harbors",2,5,"Counted","159",
  "Long Beach",1,1,"Counted","1,477","Long Beach",1,2,"Partial","494*","Long Beach",1,3,"Borrowed","borrow","Long Beach",1,4,"Borrowed","borrow","Long Beach",1,5,"Partial","48*")
stfill<-c(Counted=NAVY,Partial=AMBER,Borrowed="#EBD3D4",Closed="#EEF0F1")
sttxt <-c(Counted="white",Partial="white",Borrowed=RED,Closed="#B5BDC3")
cell$fill<-stfill[cell$status]; cell$tcol<-sttxt[cell$status]
barr <- tribble(~d,~from_y,~to_y, 2,4,2, 3,4,1, 4,3,1, 4,3,2)
p_fig18 <- ggplot() +
  geom_rect(data=cell, aes(xmin=d-0.46,xmax=d+0.46,ymin=by-0.42,ymax=by+0.42,fill=fill),color="white",linewidth=1.4) +
  geom_text(data=cell, aes(x=d,y=by,label=lab,color=tcol),fontface="bold",size=3.5) +
  geom_curve(data=barr, aes(x=d-0.30,xend=d-0.05,y=from_y-0.42,yend=to_y+0.42),curvature=-0.25,
             arrow=arrow(length=unit(0.16,"cm"),type="closed"),color=RED,linewidth=0.7) +
  scale_fill_identity() + scale_color_identity() +
  scale_x_continuous(breaks=1:5,labels=days,position="top") +
  scale_y_continuous(breaks=1:4,labels=c("Long Beach","Twin Harbors","Mocrocks","Copalis"),limits=c(0.3,4.9)) +
  annotate("text",x=0.05,y=4.75,label="NORTH / donors (counted every open day)",hjust=0,size=3.2,color=NAVY,fontface="bold") +
  annotate("text",x=0.05,y=2.66,label="SOUTH / targets (borrow on uncounted days)",hjust=0,size=3.2,color=RED,fontface="bold") +
  annotate("segment",x=0.4,xend=5.6,y=2.52,yend=2.52,linetype="dotted",color=GREY) +
  labs(title="A tide series, and how Long Beach and Twin Harbors are filled in",
       subtitle="Real 2024-25 opening series (values are digger-trips). Red arrows show a beach-day borrowed from the donor counted that day; * = only one Long Beach half walked",
       x=NULL,y=NULL,
       caption="Counted = full census   |   Partial = one Long Beach half   |   Borrowed = imputed from donor   |   blank = beach closed that day") +
  theme_minimal(base_size=13) +
  theme(panel.grid=element_blank(),axis.text.y=element_text(face="bold",size=11,color=INK),
        axis.text.x=element_text(size=10,color=INK,lineheight=0.9),
        plot.title=element_text(face="bold",size=15.5),plot.subtitle=element_text(size=10,color="grey30",margin=margin(b=8)),
        plot.caption=element_text(size=9,color="grey35",hjust=0),plot.margin=margin(10,16,10,10))
print(p_fig18); save_plot(p_fig18, "schematics", "fig18_tide_series_imputation.png", w = 13, h = 6.6)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 21: FIGURE 19 — WHY THE ESTIMATE BORROWS FROM BOTH COPALIS AND MOCROCKS
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=13, fig.height=6.4}
# The joint-donor rule (Section 4.5): X_t = X_r*(D1_t+D2_t)/(D1_r+D2_r). Summing the two donors
# before forming the ratio averages out each beach's own noise. Left: worked arithmetic. Right:
# empirical single-day imputation sd by donor set (imputation_pair_errors.csv).
lab_single <- paste0("SINGLE DONOR  (Long Beach not counted today)\n\n",
  "     LB_today  =  LB_ref  x  ( Copalis_today / Copalis_ref )\n\n",
  "     reference day:  LB = 1,200   Copalis = 1,600\n     today:          Copalis = 2,400\n\n",
  "     LB_today = 1,200 x (2,400 / 1,600) = 1,200 x 1.50 = 1,800")
lab_joint <- paste0("JOINT DONOR  (both northern beaches counted)\n\n",
  "     LB_today = LB_ref x ( Cop_today + Moc_today )/( Cop_ref + Moc_ref )\n\n",
  "     reference: LB=1,200  Cop=1,600  Moc=1,400\n     today:     Cop=2,400  Moc=2,000\n\n",
  "     LB_today = 1,200 x (4,400 / 3,000) = 1,200 x 1.47 = 1,760")
left <- ggplot() +
  geom_rect(aes(xmin=0,xmax=10,ymin=5.3,ymax=9.7),fill="#EEF3F6",color=NA) +
  geom_rect(aes(xmin=0,xmax=0.16,ymin=5.3,ymax=9.7),fill=BLUE) +
  geom_text(aes(x=0.35,y=7.5,label=lab_single),hjust=0,vjust=0.5,size=3.5,lineheight=1.15,color=INK) +
  geom_rect(aes(xmin=0,xmax=10,ymin=0.3,ymax=4.7),fill="#F0EEF6",color=NA) +
  geom_rect(aes(xmin=0,xmax=0.16,ymin=0.3,ymax=4.7),fill="#5B57A6") +
  geom_text(aes(x=0.35,y=2.5,label=lab_joint),hjust=0,vjust=0.5,size=3.5,lineheight=1.15,color=INK) +
  coord_cartesian(xlim=c(0,10),ylim=c(0,10)) + theme_void()
lbpe <- readr::read_csv(file.path(data_dir,"imputation_pair_errors.csv"),show_col_types=FALSE) %>%
  filter(target=="Long Beach") %>%
  mutate(sdlog=as.numeric(`sd of log error`),
         donor=recode(donor,"Copalis+Mocrocks"="Copalis + Mocrocks\n(joint)"),
         grp=ifelse(grepl("joint",donor),"joint","single"))
bars <- ggplot(lbpe, aes(x=reorder(donor,sdlog),y=sdlog,fill=grp)) +
  geom_col(width=0.66) +
  geom_text(aes(label=sprintf("%.2f",sdlog)),vjust=-0.4,size=3.6,fontface="bold",color=INK) +
  scale_fill_manual(values=c(single=GREY,joint=GREEN),guide="none") +
  scale_y_continuous(limits=c(0,0.75),expand=expansion(mult=c(0,0.08))) +
  labs(x=NULL,y="sd of log error (imputed vs actual)",
       title="Averaging two donors cuts the error",subtitle="Long Beach single-day imputation error, by donor set") +
  theme_minimal(base_size=12) +
  theme(panel.grid.major.x=element_blank(),panel.grid.minor=element_blank(),
        axis.text.x=element_text(face="bold",size=9.5,color=INK,lineheight=0.9),
        plot.title=element_text(face="bold",size=12.5),plot.subtitle=element_text(size=9.5,color="grey35"))
p_fig19 <- (left | bars) + plot_layout(widths=c(1.35,1)) +
  plot_annotation(title="Why the estimate borrows from BOTH Copalis and Mocrocks, not one",
    subtitle="Summing the two donors before forming the day-over-day ratio averages out each beach's own noise, leaving the shared coastwide signal",
    theme=theme(plot.title=element_text(face="bold",size=15.5),plot.subtitle=element_text(size=10.5,color="grey30")))
print(p_fig19); save_plot(p_fig19, "schematics", "fig19_joint_donor.png", w = 13, h = 6.4)
```

# ═══════════════════════════════════════════════════════════════════════════════
# SECTION 22: FIGURE 20 — LONG BEACH IS ESTIMATED THROUGH THREE STACKED ASSUMPTIONS
# ═══════════════════════════════════════════════════════════════════════════════

```{r fig.width=12, fig.height=7}
# Long Beach anatomy (Section 5.4): 7 segments walked in a north-4 and a south-3 count, plus the
# CF-from-one-end assumption and the two imputation layers.
seg <- tibble(id=c("12600","12500","12400","12300","11200","11104","11101"),
  name=c("Seaview","Long Beach","Cranberry","Klipsan","Ocean Park","","Oysterville"), x=1:7,
  half=c("S","S","S","N","N","N","N"))
seg$fill <- ifelse(seg$half=="N", BLUE, AMBER)
p_fig20 <- ggplot() +
  geom_rect(data=seg, aes(xmin=x-0.46,xmax=x+0.46,ymin=5.2,ymax=6.4,fill=fill),color="white",linewidth=1.1) +
  geom_text(data=seg, aes(x=x,y=5.8,label=id),color="white",fontface="bold",size=3.5) +
  geom_text(data=seg, aes(x=x,y=4.95,label=name),size=3.1,color=INK) +
  annotate("segment",x=0.5,xend=7.5,y=6.7,yend=6.7,arrow=arrow(length=unit(0.15,"cm"),ends="both"),color=GREY) +
  annotate("text",x=0.6,y=6.95,label="SOUTH (Beard's Hollow)",hjust=0,size=3,color=GREY) +
  annotate("text",x=7.4,y=6.95,label="NORTH (Leadbetter Pt.)",hjust=1,size=3,color=GREY) +
  annotate("segment",x=1,xend=3,y=4.5,yend=4.5,color=AMBER,linewidth=1.3) +
  annotate("text",x=2,y=4.2,label="South 3-segment count",size=3.2,color=AMBER,fontface="bold") +
  annotate("segment",x=4,xend=7,y=4.5,yend=4.5,color=BLUE,linewidth=1.3) +
  annotate("text",x=5.5,y=4.2,label="North 4-segment count",size=3.2,color=BLUE,fontface="bold") +
  geom_rect(aes(xmin=0.5,xmax=7.5,ymin=2.7,ymax=3.6),fill=RED,alpha=0.13) +
  annotate("text",x=0.7,y=3.15,hjust=0,size=3.25,color=RED,fontface="bold",
    label="Assumption 1.  The egress survey that sets the correction factor has only ever\ncovered ONE end (Oysterville-north or Seaview-south), yet CF is applied to the whole beach") +
  geom_rect(aes(xmin=0.5,xmax=7.5,ymin=1.5,ymax=2.4),fill=AMBER,alpha=0.16) +
  annotate("text",x=0.7,y=1.95,hjust=0,size=3.25,color="#9A6B1E",fontface="bold",
    label="Assumption 2.  On days only one half is walked, the other half is imputed:\nLB = X + (a/b) x X   (a, b = the two halves on a reference day),  sd of log error 0.281") +
  geom_rect(aes(xmin=0.5,xmax=7.5,ymin=0.3,ymax=1.2),fill=BLUE,alpha=0.13) +
  annotate("text",x=0.7,y=0.75,hjust=0,size=3.25,color=NAVY,fontface="bold",
    label="Assumption 3.  On fully uncounted days the WHOLE beach is borrowed from Copalis/Mocrocks:\nX_t = X_r x (D_t / D_r) for effort,  CPUE borrowed too,  sd of log error 0.541") +
  scale_fill_identity() + coord_cartesian(xlim=c(0.3,7.7),ylim=c(0.1,7.2)) +
  labs(title="Long Beach is estimated through three stacked assumptions",
       subtitle="Seven catch-record segments, walked in two halves; in 2025-26 only 13.8% of its effort rests on a full beach count") +
  theme_void() +
  theme(plot.title=element_text(face="bold",size=15.5),
        plot.subtitle=element_text(size=10.5,color="grey30",margin=margin(b=6)),plot.margin=margin(10,14,10,10))
print(p_fig20); save_plot(p_fig20, "schematics", "fig20_longbeach_anatomy.png", w = 12, h = 7)
```

```{r}
cat("\nPipeline complete. Figures written under:", folder_name, "\n")
```
