---
title: "Razor Clam Recreational Harvest: Extraction and Effort-Expansion Recompute"
author: "WDFW Razor Clam Program"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    df_print: paged
params:
  db_path: "C:/Data/Razor Clam Recreational Harvest Database iForms.accdb"
  out_dir: "./output"
  # --- impossible-value thresholds (set a cap to Inf to disable it) ---
  max_clams_per_interview: 200      # WA limit is 15 clams/digger; a group rarely exceeds ~10 diggers
  max_people_per_interview: 60
  max_diggers_per_interview: 60
  length_valid_min: 50              # mm; nullify shell lengths outside [min, max]
  length_valid_max: 200
  # --- correction factor review band (events outside are flagged, not dropped, unless drop_null_cf_events) ---
  cf_flag_min: 0.10
  cf_flag_max: 1.00
  # --- behavior switches ---
  impossible_value_action: "nullify"   # "nullify" = set bad value to NA, keep the row; "drop" = remove the row
  dedup_exact_census: true             # TRUE removes exact-duplicate census rows (same EventID, LocationID,
                                       #   NPeople, NLanterns, NCars); these are data-entry duplicates, not replicates
  count_time_col: ""                   # name of a per-count time column in tbl_DiggerCounts, if one exists; when set
                                       #   it is added to the duplicate key so counts at different times stay distinct.
                                       #   No such column exists in the database as of this writing (see protocol).
  remove_miswritten_census: true       # TRUE removes census rows where NCars >= NPeople * census_max_car_person_ratio,
                                       #   i.e. per-group records mis-entered as censuses (only event 509 in this data)
  census_max_car_person_ratio: 2       # a real census has more people than cars; a ratio this high is impossible
  multi_count_handling: "sum"          # how to combine REMAINING (non-duplicate) multiple census rows at the
                                       #   SAME (EventID, LocationID):
                                       #   "sum"  = reproduce WDFW's stored pipeline
                                       #   "mean" = treat them as replicate counts of one segment (CF-dependent; see protocol)
  drop_null_cf_events: false           # TRUE removes events with no CorrectionFactor (their harvest is otherwise NA)
  ppl_per_car_default: 2.67            # historical interview-derived average people-per-car; fallback in CalcPeople
  season_cut_month: 7                  # months >= this start a new season-year (razor clam season is fall->spring)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Purpose

This document extracts the **five raw tables** from the WDFW Razor Clam Recreational
Harvest Access database and recomputes catch-per-unit-effort (CPUE), expanded effort,
and total harvest **directly from the raw data**, using the agency's documented
effort-expansion formulas (see the companion `Effort_Expansion_Protocol.md`).

It deliberately does **not** read the database's stored derived tables
(`Spreadsheet`, `CreelSumm`, `Creel Summary by Area`, `Wastage Summary`). Those are
recomputed here from source because (a) they can be stale relative to the raw tables
and (b) their sum columns are inflated by an internal join in the summary-build query
(every creel sum is multiplied by the number of digger-count rows for the event; the
ratio used for CPUE survives, but the totals are not usable as counts).

Every row that is dropped, nullified, or flagged is written to an **audit log** so the
estimates remain defensible.

> The formula chain implemented below was verified offline: applied to the database's
> own inputs it reproduces the stored `Spreadsheet` values for `CalcPeople`,
> `EstCorEffort`, and `TotHarvest` to floating-point precision.

# Requirements and connection notes

- 64-bit R must be paired with the **64-bit** Microsoft Access Database Engine
  (ACE) driver, or 32-bit R with the 32-bit driver. A bitness mismatch is the most
  common `IM002 ... data source name not found` failure.
- The `.accdb` is ACE14 (Access 2010+); the modern driver name is
  `Microsoft Access Driver (*.mdb, *.accdb)`.

```{r libs}
need <- c("DBI", "odbc", "dplyr", "tidyr", "stringr", "lubridate", "readr", "tibble")
miss <- need[!vapply(need, requireNamespace, logical(1), quietly = TRUE)]
if (length(miss)) stop("Install missing packages: ", paste(miss, collapse = ", "))
suppressPackageStartupMessages(lapply(need, library, character.only = TRUE))

dir.create(params$out_dir, showWarnings = FALSE, recursive = TRUE)

# Robust date parser: passes through native Date/POSIXct (the usual ODBC result) and
# falls back to parsing character dates in ISO 'Y-m-d' or US 'm/d/y' form, with or
# without a time component, returning NA for anything unparseable.
as_date_safe <- function(x) {
  if (inherits(x, "Date"))    return(x)
  if (inherits(x, "POSIXct")) return(as.Date(x))
  as.Date(lubridate::parse_date_time(
    as.character(x), orders = c("Ymd HMS", "Ymd", "mdy HMS", "mdy"), quiet = TRUE))
}
```

```{r audit-helper}
# Accumulating audit log. Every cleaning action appends one row.
.audit <- list()
log_audit <- function(table, step, action, reason, n_rows, detail = "") {
  .audit[[length(.audit) + 1]] <<- tibble(
    table = table, step = step, action = action, reason = reason,
    n_rows = as.integer(n_rows), detail = detail,
    timestamp = format(Sys.time(), "%Y-%m-%d %H:%M:%S")
  )
}
audit_table <- function() if (length(.audit)) bind_rows(.audit) else tibble()
```

# 1. Extract the raw tables

Explicit column lists avoid the empty `Field1`-`Field15` import artifacts and the Memo
columns in `tbl_events`. `lut_Beach` is read as a small reference for beach identity and
`LocationID` validation (it is a lookup, not one of the five data tables).

```{r connect-extract}
conn_str <- sprintf(
  "Driver={Microsoft Access Driver (*.mdb, *.accdb)};DBQ=%s;Uid=Admin;ReadOnly=1;",
  params$db_path
)
con <- dbConnect(odbc::odbc(), .connection_string = conn_str)

q <- function(sql) as_tibble(dbGetQuery(con, sql))

events_raw <- q("SELECT EventID, BeachName, [Date], [Low Tide], TideHeight,
                        CorrectionFactor, Sampler1
                 FROM [tbl_events]")
creel_raw  <- q("SELECT EventID, LocationID, CreelNPeople, CreelNDiggers, CreelNClams,
                        CreelNLanterns, CreelNCars
                 FROM [tbl_Creel]")
dc_cols <- "EventID, LocationID, NPeople, NLanterns, NCars"
if (nzchar(params$count_time_col)) dc_cols <- paste0(dc_cols, ", [", params$count_time_col, "]")
dc_raw     <- q(paste0("SELECT ", dc_cols, " FROM [tbl_DiggerCounts]"))
wast_raw   <- q("SELECT EventID, LocationID, NHoles, NClams, NMorts
                 FROM [tbl_Wastage]")
len_raw    <- q("SELECT EventID, LocationID, [Length] AS Length
                 FROM [TBL Length]")
lut_beach  <- q("SELECT LocationID, BeachName FROM [lut_Beach]")

dbDisconnect(con)

# raw counts -> audit
for (nm in c("tbl_events","tbl_Creel","tbl_DiggerCounts","tbl_Wastage","TBL Length")) {
  obj <- switch(nm, "tbl_events"=events_raw, "tbl_Creel"=creel_raw,
                "tbl_DiggerCounts"=dc_raw, "tbl_Wastage"=wast_raw, "TBL Length"=len_raw)
  log_audit(nm, "extract", "read", "raw rows pulled from Access", nrow(obj))
}

# normalize types
to_num <- function(x) suppressWarnings(as.numeric(x))
events_raw <- events_raw %>%
  mutate(EventID = as.integer(EventID),
         Date = as_date_safe(Date),
         CorrectionFactor = to_num(CorrectionFactor),
         TideHeight = to_num(TideHeight))
creel_raw <- creel_raw %>% mutate(EventID = as.integer(EventID),
         across(starts_with("Creel"), to_num), LocationID = as.character(LocationID))
dc_raw <- dc_raw %>% mutate(EventID = as.integer(EventID),
         across(c(NPeople, NLanterns, NCars), to_num), LocationID = as.character(LocationID))
wast_raw <- wast_raw %>% mutate(EventID = as.integer(EventID),
         across(c(NHoles, NClams, NMorts), to_num), LocationID = as.character(LocationID))
len_raw <- len_raw %>% mutate(EventID = as.integer(EventID),
         Length = to_num(Length), LocationID = as.character(LocationID))
```

# 2. Cleaning layer (audited)

## 2.1 Canonical beach identity

The event's stored `BeachName` is not normalized ("LongBeach" / "Long Beach" /
"Longbeach" are one beach). We map to the `lut_Beach` canonical spelling. For
child-level QC we also validate `LocationID` against `lut_Beach`.

```{r beach-clean}
canon_beach <- function(x) {
  k <- str_to_lower(str_replace_all(as.character(x), "\\s+", ""))
  dplyr::recode(k,
    "longbeach"   = "LongBeach",
    "twinharbors" = "Twin Harbors",
    "copalis"     = "Copalis",
    "mocrocks"    = "Mocrocks",
    "kalaloch"    = "Kalaloch",
    .default = NA_character_)
}

events <- events_raw %>% mutate(beach = canon_beach(BeachName))
n_unmapped <- sum(is.na(events$beach))
if (n_unmapped) log_audit("tbl_events", "normalize_beach", "flag",
  "BeachName not mappable to a canonical beach", n_unmapped,
  paste(unique(events_raw$BeachName[is.na(events$beach)]), collapse = "; "))

valid_loc <- unique(as.character(lut_beach$LocationID))
```

## 2.2 Correction factor review

`CorrectionFactor` drives effort expansion; out-of-band or missing values are flagged
(and optionally the events dropped). A value > 1 is implausible if CF is the fraction of
diggers present at count time.

```{r cf-clean}
cf_bad   <- with(events, !is.na(CorrectionFactor) &
                   (CorrectionFactor < params$cf_flag_min | CorrectionFactor > params$cf_flag_max))
cf_null  <- is.na(events$CorrectionFactor)
if (any(cf_bad)) log_audit("tbl_events", "cf_review", "flag",
  sprintf("CorrectionFactor outside [%.2f, %.2f]", params$cf_flag_min, params$cf_flag_max),
  sum(cf_bad), paste(sort(unique(events$EventID[cf_bad])), collapse = ","))
if (any(cf_null)) log_audit("tbl_events", "cf_review", "flag",
  "CorrectionFactor missing (effort/harvest will be NA)", sum(cf_null),
  paste(sort(events$EventID[cf_null]), collapse = ","))

if (isTRUE(params$drop_null_cf_events) && any(cf_null)) {
  log_audit("tbl_events", "cf_review", "drop", "drop_null_cf_events = TRUE", sum(cf_null))
  events <- events %>% filter(!is.na(CorrectionFactor))
}
valid_eid <- unique(events$EventID)
```

## 2.3 Generic child cleaning: null/orphan keys, location validation

```{r child-keys}
clean_child_keys <- function(df, tbl) {
  n0 <- nrow(df)
  # null EventID -> unlinkable -> drop
  nnull <- sum(is.na(df$EventID))
  if (nnull) { log_audit(tbl, "keys", "drop", "EventID is NULL (unlinkable)", nnull)
               df <- df %>% filter(!is.na(EventID)) }
  # orphan EventID -> no parent in tbl_events -> drop
  orph <- !(df$EventID %in% valid_eid)
  if (any(orph)) { log_audit(tbl, "keys", "drop", "EventID has no parent in tbl_events",
                             sum(orph), paste(sort(unique(df$EventID[orph])), collapse = ","))
                   df <- df %>% filter(EventID %in% valid_eid) }
  # bad LocationID -> flag only (event-level math keys on EventID, not LocationID)
  if ("LocationID" %in% names(df)) {
    badloc <- !(df$LocationID %in% valid_loc) & !is.na(df$LocationID) & df$LocationID != ""
    if (any(badloc)) log_audit(tbl, "keys", "flag", "LocationID not in lut_Beach",
                               sum(badloc),
                               paste(utils::head(sort(unique(df$LocationID[badloc])), 20), collapse = ","))
  }
  df
}

creel <- clean_child_keys(creel_raw, "tbl_Creel")
dc    <- clean_child_keys(dc_raw,    "tbl_DiggerCounts")
wast  <- clean_child_keys(wast_raw,  "tbl_Wastage")
len   <- clean_child_keys(len_raw,   "TBL Length")
```

## 2.3b Remove exact-duplicate census rows

Some events have the same instantaneous count entered more than once (in a few cases an
entire event's census was duplicated). These are data-entry duplicates, not replicate
counts, and summing them double-counts effort. We drop exact duplicates on
`(EventID, LocationID, NPeople, NLanterns, NCars)`, keeping one. Note: if replicate
counts taken at different times could legitimately share identical values, add a time or
comment field to this key; here counts are treated as duplicates when location and all
count values match.

```{r dedup}
if (isTRUE(params$dedup_exact_census)) {
  key <- c("EventID", "LocationID", "NPeople", "NLanterns", "NCars")
  if (nzchar(params$count_time_col) && params$count_time_col %in% names(dc))
    key <- c(key, params$count_time_col)
  before <- nrow(dc)
  dc <- dc %>% distinct(across(all_of(key)), .keep_all = TRUE)
  removed <- before - nrow(dc)
  if (removed) log_audit("tbl_DiggerCounts", "dedup", "drop",
    paste0("exact-duplicate census row on (", paste(key, collapse = ", "), ")"),
    removed, "kept one row per identical group")
}
```

## 2.3c Repair mis-entered census rows

A census row records people, cars, and lanterns on a beach segment, so it must have more
people than cars. A handful of rows (in this data, only event 509) are per-group records
typed into the census table by mistake, identifiable by an impossible car-to-people ratio
(`NCars` set to `NPeople` times the bag limit). These are removed so they do not inflate
the census. Note: removing them does not by itself produce a harvest estimate for an event
that also lacks creel interviews (event 509 has none, so it remains without a CPUE and
yields no harvest); it only cleans the effort census.

```{r repair-census}
if (isTRUE(params$remove_miswritten_census)) {
  bad <- which(!is.na(dc$NPeople) & !is.na(dc$NCars) & dc$NPeople > 0 &
               dc$NCars >= dc$NPeople * params$census_max_car_person_ratio)
  if (length(bad)) {
    log_audit("tbl_DiggerCounts", "repair", "drop",
      sprintf("NCars >= NPeople * %s (per-group records mis-entered as census)",
              params$census_max_car_person_ratio),
      length(bad), paste0("events ", paste(sort(unique(dc$EventID[bad])), collapse = ",")))
    dc <- dc[-bad, , drop = FALSE]
  }
}
```

## 2.4 Impossible numeric values

`CreelNCars` is renamed to `ppl_per_car` (it stores people-per-car, not a vehicle
count). Values above the configured caps are nullified (default) or their rows dropped.

```{r impossible}
creel <- creel %>% rename(ppl_per_car = CreelNCars)

apply_cap <- function(df, col, cap, tbl) {
  if (!is.finite(cap)) return(df)
  hit <- which(!is.na(df[[col]]) & df[[col]] > cap)
  if (!length(hit)) return(df)
  vals <- paste(utils::head(sort(df[[col]][hit], decreasing = TRUE), 8), collapse = ",")
  if (params$impossible_value_action == "drop") {
    log_audit(tbl, "impossible", "drop", sprintf("%s > %s (row removed)", col, cap),
              length(hit), vals)
    df <- df[-hit, , drop = FALSE]
  } else {
    log_audit(tbl, "impossible", "nullify", sprintf("%s > %s (set to NA)", col, cap),
              length(hit), vals)
    df[[col]][hit] <- NA_real_
  }
  df
}
creel <- creel %>%
  apply_cap("CreelNClams",   params$max_clams_per_interview,   "tbl_Creel") %>%
  apply_cap("CreelNPeople",  params$max_people_per_interview,  "tbl_Creel") %>%
  apply_cap("CreelNDiggers", params$max_diggers_per_interview, "tbl_Creel")

# Length: nullify outside the biologically plausible band (length is not used in the
# harvest calc but is cleaned and exported for length-frequency work)
lbad <- which(!is.na(len$Length) &
                (len$Length < params$length_valid_min | len$Length > params$length_valid_max))
if (length(lbad)) {
  log_audit("TBL Length", "impossible", "nullify",
            sprintf("Length outside [%g, %g] mm", params$length_valid_min, params$length_valid_max),
            length(lbad), paste(sort(unique(len$Length[lbad])), collapse = ","))
  len$Length[lbad] <- NA_real_
}
```

# 3. Recompute the effort-expansion chain

All event-level rates come from the **cleaned raw creel**, not the stored summaries.

## 3.1 Event-level creel rates

```{r creel-rates}
creel_event <- creel %>%
  group_by(EventID) %>%
  summarise(
    SumClams    = sum(CreelNClams,    na.rm = TRUE),
    SumDiggers  = sum(CreelNDiggers,  na.rm = TRUE),
    SumPeople   = sum(CreelNPeople,   na.rm = TRUE),
    SumLanterns = sum(CreelNLanterns, na.rm = TRUE),
    AvgPplPerCar = mean(ppl_per_car,  na.rm = TRUE),  # NaN if no car data this event
    n_interviews = n(),
    .groups = "drop"
  ) %>%
  mutate(
    CPUE            = ifelse(SumDiggers  > 0, SumClams   / SumDiggers,  NA_real_),
    DiggersPeople   = ifelse(SumPeople   > 0, SumDiggers / SumPeople,   NA_real_),
    DiggersLanterns = ifelse(SumLanterns > 0, SumDiggers / SumLanterns, NA_real_),
    AvgPplPerCar    = ifelse(is.nan(AvgPplPerCar), NA_real_, AvgPplPerCar)
  )
```

## 3.2 Per-census expansion (Spreadsheet grain)

Each `tbl_DiggerCounts` row is expanded with the hierarchical `CalcPeople` rule, then
multiplied through to `TotHarvest`. This reproduces the database's `Spreadsheet` rows.

```{r expand}
working <- dc %>%
  left_join(events %>% select(EventID, beach, Date, CorrectionFactor), by = "EventID") %>%
  left_join(creel_event %>% select(EventID, CPUE, DiggersPeople, DiggersLanterns, AvgPplPerCar),
            by = "EventID") %>%
  mutate(
    CalcPeople = dplyr::case_when(
      !is.na(NPeople)   ~ NPeople,
      !is.na(NCars)     ~ NCars * AvgPplPerCar,
      !is.na(NLanterns) ~ NLanterns * DiggersLanterns,
      TRUE              ~ NCars * params$ppl_per_car_default
    ),
    EstCorEffort = CalcPeople * DiggersPeople / CorrectionFactor,
    TotHarvest   = EstCorEffort * CPUE
  )

# transparency: report multiple census rows at the same (EventID, LocationID)
multi <- working %>% count(EventID, LocationID) %>% filter(n > 1)
if (nrow(multi)) log_audit("tbl_DiggerCounts", "grain", "flag",
  sprintf("multiple census rows at same (EventID, LocationID); handling = '%s'", params$multi_count_handling),
  sum(multi$n), paste0(nrow(multi), " segment-pairs; max ", max(multi$n), " rows; events ",
                        paste(utils::head(sort(unique(multi$EventID)), 10), collapse = ",")))
```

## 3.3 Aggregate to event and beach-season

`multi_count_handling` decides how replicate census rows at one segment combine:
`"sum"` reproduces WDFW; `"mean"` averages replicate counts of a segment before
summing across segments (recommended; the correction factor already accounts for
within-day turnover).

```{r aggregate}
if (params$multi_count_handling == "mean") {
  seg <- working %>% group_by(EventID, LocationID) %>%
    summarise(EstCorEffort = mean(EstCorEffort, na.rm = TRUE),
              TotHarvest   = mean(TotHarvest,   na.rm = TRUE), .groups = "drop")
} else {
  seg <- working %>% select(EventID, LocationID, EstCorEffort, TotHarvest)
}

wastage_event <- wast %>% group_by(EventID) %>%
  summarise(SumMorts = sum(NMorts, na.rm = TRUE),
            SumHoles = sum(NHoles, na.rm = TRUE), .groups = "drop") %>%
  mutate(WastageRate = ifelse(SumHoles > 0, SumMorts / SumHoles, NA_real_))

event_harvest <- seg %>%
  group_by(EventID) %>%
  summarise(EstCorEffort = sum(EstCorEffort, na.rm = TRUE),
            TotHarvest   = sum(TotHarvest,   na.rm = TRUE), .groups = "drop") %>%
  left_join(events %>% select(EventID, beach, Date, CorrectionFactor), by = "EventID") %>%
  left_join(creel_event %>% select(EventID, CPUE, SumPeople, n_interviews), by = "EventID") %>%
  left_join(wastage_event %>% select(EventID, WastageRate), by = "EventID") %>%
  mutate(
    Wastage      = TotHarvest * WastageRate,
    PctInterviewed = ifelse(EstCorEffort > 0, SumPeople / EstCorEffort, NA_real_),
    year  = lubridate::year(Date),
    month = lubridate::month(Date),
    season_year  = ifelse(month >= params$season_cut_month, year, year - 1L),
    season_label = paste0(season_year, "-", substr(as.character(season_year + 1L), 3, 4))
  )

beach_season <- event_harvest %>%
  filter(!is.na(beach)) %>%
  group_by(beach, season_year, season_label) %>%
  summarise(
    n_events       = n(),
    total_effort   = sum(EstCorEffort, na.rm = TRUE),
    total_harvest  = sum(TotHarvest,   na.rm = TRUE),
    total_wastage  = sum(Wastage,      na.rm = TRUE),
    mean_cpue      = mean(CPUE,        na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(beach, season_year)
```

# 4. Outputs and audit

```{r write}
w <- function(x, f) readr::write_csv(x, file.path(params$out_dir, f))
w(events,         "clean_events.csv")
w(creel,          "clean_creel.csv")
w(dc,             "clean_diggercounts.csv")
w(wast,           "clean_wastage.csv")
w(len,            "clean_length.csv")
w(creel_event,    "creel_event_rates.csv")
w(working,        "working_spreadsheet_grain.csv")
w(event_harvest,  "harvest_by_event.csv")
w(beach_season,   "harvest_by_beach_season.csv")

audit <- audit_table()
w(audit, "audit_log.csv")
```

```{r audit-summary}
cat("Rows dropped / nullified / flagged by table and action:\n")
audit %>% group_by(table, action) %>%
  summarise(events_or_rows = sum(n_rows), n_actions = n(), .groups = "drop") %>%
  arrange(table, action) %>% as.data.frame() %>% print(row.names = FALSE)
```

```{r preview}
cat("\nHarvest by beach and season (head):\n")
head(beach_season, 15)
```

# 5. Notes and known issues carried into the estimate

- **Derived tables avoided.** All aggregates are recomputed from raw; the database's
  stored summary tables are not used (stale, and their totals are inflated by a join).
- **`CreelNCars` = people per car**, renamed `ppl_per_car`. Do not read it as a vehicle count.
- **Exact-duplicate censuses removed.** Some events had a census entered twice (in a few
  cases the whole event was duplicated, e.g. Twin Harbors events 2115, 2118, 2123, 2668);
  `dedup_exact_census = TRUE` drops these. This alone corrects an ~16% overstatement of the
  Twin Harbors 2022-23 season. The audit log records how many rows were removed.
- **Multiple census rows per segment (after dedup).** A few events still have genuine
  distinct counts at one segment (different values, likely different times). With
  `multi_count_handling = "sum"` these are summed (WDFW behavior); `"mean"` averages them.
  After deduplication this choice affects only ~0.25% of the all-time total, concentrated in
  a handful of seasons, and is genuinely correction-factor-dependent. The mis-entered census
  rows in event 509 (Long Beach) are now removed by `remove_miswritten_census`; that event
  still yields no harvest because it has no creel interviews, so its census is corrected but
  no CPUE exists to expand.
- **Missing correction factors** leave an event's effort and harvest as `NA` unless
  `drop_null_cf_events = TRUE` or a CF is imputed; the audit lists the affected events.
- **Length** is cleaned and exported but is not part of the harvest calc; only ~16% of
  events have length data.
- All thresholds are `params` at the top; change them and re-knit. The audit log records
  exactly what each setting removed or altered.
```
